---
title: |
  Automated assessment of residual plots with computer vision models
type: Draft paper (Incomplete)
author:
  - name: Weihao Li
    affil: a
    email: weihao.li@monash.edu
  - name: Dianne Cook
    affil: a
    email: dicook@monash.edu
  - name: Emi Tanaka
    affil: b, c
    email: emi.tanaka@anu.edu.au
  - name: Susan VanderPlas
    affil: d
    email: susan.vanderplas@unl.edu
  - name: Klaus Ackermann
    affil: a
    email: Klaus.Ackermann@monash.edu
affiliation:
  - num: a
    address: |
      Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
  - num: b
    address: |
      Biological Data Science Institute, Australian National University, Acton, ACT, Australia
  - num: c
    address: |
      Research School of Finance, Actuarial Studies and Statistics, Australian National University, Acton, ACT, Australia
  - num: d
    address: |
      Department of Statistics, University of Nebraska, Lincoln, Nebraska, USA
bibliography: ref.bib
abstract: |
  TBD.
keywords: |
  TBD
header-includes: |
  \usepackage{lscape}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
output: rticles::tf_article
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  echo = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "100%",
  fig.align = "center")
```

```{r}
# Visual inference models and p-value calculation
# remotes::install_github("TengMCing/visage)

library(tidyverse)
library(visage)
library(glue)

# To control the simulation in this file
set.seed(10086)
```


# Introduction

Plotting residuals is commonly regarded as a standard practice in linear regression diagnostics \citep[see][]{cook1982residuals, belsley1980regression}. This visual assessment plays a crucial role in identifying deviations from model assumptions, such as linearity, homoscedasticity, and normality. It also helps in understanding the goodness of fit and various characteristics of the model.

Generating a residual plot in most statistical software is often as straightforward as executing a line of code or clicking a button. However, accurately interpreting a residual plot can be challenging. A residual plot can exhibit various visual features, but it is crucial to recognize that some may arise from the characteristics of predictors and the inherent randomness of the error, rather than indicating a violation of model assumptions [@li2023plot]. Consider Figure \ref{fig:false-finding} as an example, the residuals display a triangular shape pointing to the left. While this might suggest heteroskedasticity, it is important to avoid over-interpreting this visual pattern. In this case, the fitted regression model is correctly specified, and the triangular shape is actually a result of the skewed distribution of the predictors, rather than indicating a flaw in the model.

The concept of visual inference, as proposed by @buja2009statistical, provides an inferential framework to assess whether residual plots indeed contain visual patterns inconsistent with the model assumptions. The fundamental idea involves testing whether the true residual plot visually differs significantly from null plots, where null plots are plotted with residuals generated from the residual rotation distribution [@langsrud2005rotation], which is a distribution consistent with the null hypothesis $H_0$ that the linear regression model is correctly specified. Typically, the visual test is accomplished through the lineup protocol, where the real residual plot is embedded within a lineup alongside several null plots. If the real residual plot can be distinguished from the lineup, it provides evidence for rejecting $H_0$.

The practice of delivering a residual plot as a lineup is generally regarded as a valuable approach. Beyond its application in residual diagnostics, the lineup protocol has integrated into the analysis of diverse subjects. For instance, \cite{loy2013diagnostic, loy2014hlmdiag, loy2015you} illustrated its applicability in diagnosing hierarchical linear models. Additionally, @widen2016graphical demonstrated its utility in geographical research, while @krishnan2021hierarchical explored its effectiveness in forensic examinations.

However, as pointed out by @li2023plot, a primary limitation of the lineup protocol lies in its reliance on human judgments. Unlike conventional statistical tests that can be performed computationally in statistical software, the lineup protocol requires human evaluation of images. This characteristic makes it less suitable for large-scale applications, given the associated high labour costs and time requirements.

There is a substantial need to develop an approach that alleviates analysts' workload by automating repetitive tasks and providing standardized results in a controlled environment. The large-scale evaluation of lineups is impractical without the use of technology and machines.

The utilization of computers to interpret data plots has a rich history, with early efforts such as "Scagnostics" by @tukey1985computer, focusing on scatter plot diagnostics. @wilkinson2005graph expanded on this work, introducing graph theoretic scagnostics, which encompassed computable measures applied to planar proximity graphs. These measures, including, but not limited to, "Outlying", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy", and "Striated" aimed to characterize outliers, shape, density, trend, coherence and other characteristics of the data. While this approach has been inspiring, there is a recognition [@buja2009statistical] that it may not capture all the necessary visual features that differentiate true residual plots from null plots. A more promising alternative entails enabling machines to learn the function for extracting visual features from residual plots. Essentially, this means empowering computers to discern the crucial visual features for residual diagnostics and determining the method to extract them. 

Modern computer vision models are well-suited for addressing this challenge. They rely on deep neural networks with convolutional layers [@fukushima1982neocognitron]. These layers leverage hierarchical patterns in data, downsizing and transforming images by summarizing information in a small space. Numerous studies have demonstrated the efficacy of convolutional layers in addressing various vision tasks, including image recognition [@rawat2017deep]. Despite the widespread use of computer vision models in fields like computer-aided diagnosis [@lee2015image], pedestrian detection [@brunetti2018computer], and facial recognition [@emami2012facial], their application in reading data plots remains limited. While some studies have explored the use of computer vision models for tasks such as reading recurrence plots for time series regression [@ojeda2020multivariate], time series classification [@chu2019automatic; @hailesilassie2019financial; @hatami2018classification; @zhang2020encoding], anomaly detection [@chen2020convolutional], and pairwise causality analysis [@singh2017deep], the application of reading residual plots with computer vision models represents a relatively new field of study.

In this paper, we develop computer vision models and integrate them into the residual plots diagnostics workflow, filling the gap of.... The paper is structured as follows: ...


```{r false-finding, fig.width = 8/5*4, fig.height = 6/4*4, fig.pos="!h", fig.cap="An example residual vs fitted values plot (red line indicates 0). The vertical spread of the data points varies with the fitted values. This often indicates the existence of heteroskedasticity. The Breusch-Pagan test rejects this residual plot at 95\\% significance level ($p\\text{-value} = 0.046$)."}
set.seed(452)
ori_x <- rand_lognormal()
mod <- heter_model(b = 0, x = closed_form(~-ori_x))
ori_dat <- mod$gen(300)

ori_dat %>%
  VI_MODEL$plot(theme = theme_light(), remove_grid_line = TRUE) +
  xlab("Fitted values") +
  ylab("Residuals")
```


# Methodology

## Model Specifications

There are various specifications of the computer vision model that can be used to assess residual plots. We discuss these specifications below focusing on two key components of the model formula: the input and the output format.


### Input Formats

Deep learning models are in general very sensitive to the input data. The quality and relevance of the input data greatly influence the model's capacity to generate insightful and meaningful results. There are several designs of the input format can be considered.

A straightforward design involves feeding a vector of residuals along with a vector of fitted values, essentially providing all the necessary information for creating a residuals vs fitted values plot. However, a drawback of this method is the dynamic input size, which changes based on the number of observations. For modern computer vision models implemented in mainstream software like TensorFlow [@abadi2016tensorflow], the input shape is typically fixed. One solution is to pad the input vectors with leading or trailing zeros when the input tensor expects longer vectors, but it may fail if the input vector surpasses the designed length. Another strategy is to summarize the residuals and fitted values separately using histograms and utilize the counts as the input. By controlling the number of bins in the histograms, it becomes possible to provide fixed-length input vectors. Still, since histograms only capture the marginal distribution of residuals and fitted values respectively, they can not be used to differentiate visual patterns with same marginal distributions but different joint distributions.

Another design involves using an image as input. The primary advantage of this design, as opposed to the vector format, is the availability of the existing and sophisticated image processing architectures developed over the years, such as the VGG16 architecture proposed in @simonyan2014very. These architectures can effectively capture and summarize spatial information from nearby pixels, which is less straightforward with vector input. The main considerations are the image resolution and the aesthetics of the residual plot. In general, higher resolution provides more information to the model but comes with the trade-off of increased complexity and greater difficulty in training. As for the aesthetics of the residual plot, a practical solution is to consistently present residual plots in the same style to the model. This implies that the model can not accept arbitrary images as input but requires the use of the same pre-processing pipeline to convert residuals and fitted values into a standardized-style residual plot.

Providing multiple residual plots to the model, such as a pair of plots, a triplet or a lineup is also a possible option. @chopra2005learning have shown that computer vision models designed for image comparison can assess whether a pair of images are similar or dissimilar. Applied to our specific problem, we can define null plots of a fitted regression model to be similar to each other, while considering true residual plots to be distinct from null plots of any fitted regression model. A triplet constitutes a set of three images, denoted as $image_1$, $image_2$ and $image_3$. It is often used to predict whether $image_2$ or $image_3$ is more similar to $image_1$, proving particularly useful for establishing rankings between samples. For this setup, we can apply the same criteria to define similarity between images. However, it is important to note that these two approaches usually require additional considerations regarding the loss function and, at times, non-standard training processes due to shared weights between different convolutional blocks.

Presenting a lineup to a model aligns closely with the lineup protocol. However, as the number of residual plots in a lineup increases, the resolution of the input image grows rapidly, posing challenges in training the model. We experimented with this approach in a pilot study, but the performance of the trained model was sub-optimal.

We did not explore all the mentioned input formats due to the significant costs tied to data preparation and model training. Taking into account the implementation cost and the need for model interpretability, we settled on the single residual plot input format.

### Output Formats

Given that the input is a single residual plot represented as a fixed-resolution image, the output from the computer vision model can take one of two forms: binary or numeric. This choice determines whether the model belongs to a classification model or a regression model. The binary outcome encoded as $0$ and $1$ could be used to represent whether the input image is a null plot, or whether the input image would be rejected in a visual test conducted by humans. Training a model following the latter option requires data from prior human subject experiments, presenting difficulties in controlling the quality of data due to variations in experimental settings across different studies. Additionally, some visual inference experiments are unrelated to linear regression models or residual plot diagnostics, resulting in a limited amount of available training data.

Alternatively, the output could be a meaningful and interpretable numerical measure useful for assessing residual plots, such as the strength of suspicious visual patterns reflecting the extent of model violations, or the difficulty index for identifying whether a residual plot has no issues. However, these numeric measures are often informally used in daily communication but are not typically formalized or rigorously defined. For the purpose of training a model, this numeric measure has to be quantifiable. 

In this study, we chose to define and use a distance measure to quantify the difference between the residual plot and the null plots. @vo2016localizing have also demonstrated that defining a proper distance between images can enhance the matching accuracy in image search compared to a binary outcome model.


## Distance Between Residual Plots

In a visual test, the observer will be asked to choose one or more plots that stand out as most distinct from others in a given lineup. To develop a computer vision model for assessing residual plots within the visual inference framework, it is important to precisely define a numerical measure of "difference" or "distance" between plots. This distance can take the form of a basic statistical operation on pixels, such as the sum of square differences. Alternatively, it could involve established image similarity metrics like the Structural Similarity Index Measure [@wang2004image]. The challenge lies in the fact that metrics tailored for image comparison may not be suitable for evaluating data plots, where only essential plot elements require assessment [@chowdhury2018measuring]. Furthermore, scagnostics mentioned in Section \ref{introduction} could be used to construct distance metrics for residual plots comparison, but the functional form still needs to be carefully refined to accurately reflect the extent of the violations.

### Residual Distribution

The distance measure proposed in this study takes into account the fact that we tried to measure how different a residual plot is from a good residual plot, or in other words, how different a given fitted regression model is from a correctly specified model. 

For a classical normal linear regression model, residuals $\boldsymbol{e}$ are derived from the fitted values $\hat{\boldsymbol{y}}$ and observed values $\boldsymbol{y}$. Suppose the data generating process is known and the regression model is correctly specified, by the Frisch-Waugh-Lowell theorem [@frisch1933partial], residuals $\boldsymbol{e}$ can also be treated as random variables and written as a linear transformation of the error $\boldsymbol{\varepsilon}$ formulated as $\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon}$, where $\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the residual operator and an idempotent matrix, $\boldsymbol{X}$ is the design matrix, $\boldsymbol{I}_n$ is a $n$ by $n$ identity matrix, and $n$ is the number of observations.

One of the assumptions of the classical normal linear regression model is the error $\boldsymbol{\varepsilon}$ follows a multivariate normal distribution with zero mean and constant variance, i.e., $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$. It can be known that residuals $\boldsymbol{e}$ also follow a certain probability distribution transformed from the multivariate normal distribution, which will be denoted as $Q$. This reference distribution $Q$ summarizes what good residuals should follow given the design matrix $\boldsymbol{X}$ is known and fixed.

Suppose the design matrix $\boldsymbol{X}$ has linearly independent columns, the trace of the hat matrix will equal to the number of columns in $\boldsymbol{X}$ denoted as $k$. As a result, the rank of $\boldsymbol{R}$ is $n - k$, and $Q$ is a degenerate multivariate distribution. To capture the characteristics of $Q$, such as moments, we can simulate a large numbers of $\boldsymbol{\varepsilon}$ and transform it to $\boldsymbol{e}$ to get the empirical estimates. For simplicity, in this study, we replaced the variance-covariance matrix of residuals $\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2$ with a full-rank diagonal matrix $\text{diag}(\boldsymbol{R}\sigma^2)$, where $\text{diag}(.)$ sets the non-diagonal entries of a matrix to zeros. The resulting distribution for $\boldsymbol{Q}$ is $N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\sigma^2))$.

Distribution $Q$ is derived from the correctly specified model. However, if the model is misspecified, then the actual distribution of residuals denoted as $P$, will be different from $Q$. For example, if the data generating process contains variables correlated with any column of $\boldsymbol{X}$ but missing from $\boldsymbol{X}$, causing an omitted variable problem, $P$ will be different from $Q$ because the residual operator obtained from the fitted regression model will not be the same as $\boldsymbol{R}$. Besides, if the $\boldsymbol{\varepsilon}$ follows a non-normal distribution such as a multivariate log-normal distribution, $P$ will usually be skewed and has a long tail. 

### Distance of $P$ from $Q$

Define a proper distance between distributions is usually easier than define a proper distance between data plots. Given the true residual distribution $Q$ and the reference residual distribution $P$, we used a distance measure based on Kullback-Leibler divergence [@kullback1951information] to quantify the difference between two distributions

\begin{align}
\label{eq:kl-0}
D &= \log\left(1 + D_{KL}\right), \\
\label{eq:kl-1}
D_{KL} &= \int_{\mathbb{R}^{n}}\log\frac{p(\boldsymbol{e})}{q(\boldsymbol{e})}p(\boldsymbol{e})d\boldsymbol{e},
\end{align}

\noindent where $p(.)$ and $q(.)$ are the probability density functions for distribution $P$ and distribution $Q$ respectively.

This distance measure was first proposed in @li2023plot. It was mainly designed for measuring the effect size of non-linearity and heteroskedasticity in a residual plot. @li2023plot have derived that, for a classical normal linear regression model that omits a necessary higher-order predictors $\boldsymbol{Z}$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ with $\boldsymbol{V}$ being an arbitrary symmetric positive semi-definite matrix, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))$. Note that the variance-covariance matrix is replaced with the diagonal matrix to ensure it is a full-rank matrix. 

Since both $P$ and $Q$ are adjusted to be multivariate normal distributions, equation \ref{eq:kl-1} can be further expanded to

\begin{align}
\label{eq:kl-2}
D_{KL} &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + tr(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
\end{align}

\noindent where $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$, and $\boldsymbol{W} = \boldsymbol{R}\boldsymbol{V}\boldsymbol{R}$. The assumed error variance $\sigma^2$ is set to be $tr(\boldsymbol{V})/n$, which is the expectation of the estimated variance.


### Non-normal $P$

For non-normal error $\boldsymbol{\varepsilon}$, the true residual distribution $P$ is unlikely to be a multivariate normal distribution. Thus, equation \ref{eq:kl-2} given in @li2023plot will not be applicable to models violating the normality assumption. 

To evaluate the Kullback-Leibler divergence of non-normal $P$ from $Q$, the fallback is to solve equation \ref{eq:kl-1} numerically. However, since $\boldsymbol{e}$ is a linear transformation of non-normal random variables, it is very common that the general form of $P$ is unknown, meaning that we can not easily compute $p(\boldsymbol{e})$ using a well-known probability density function. Additionally, even if $p(\boldsymbol{e})$ can be calculated for any $\boldsymbol{e} \in \mathbb{R}^n$, it will be very difficult to do numerical integration over the $n$ dimensional space, because $n$ could be potentially very large.   

In order to approximate $D_{KL}$ in a practically computable manner, the elements of $\boldsymbol{e}$ are assumed to be independent of each other. This assumption solves both of the issues mentioned above. First, we no longer need to integrate over $n$ random variables. The result of equation \ref{eq:kl-1} is now the sum of the Kullback-Leibler divergence evaluated for each individual residual thanks for the independence assumption. Second, it is not required to know the joint probability density $p(\boldsymbol{e})$ any more. Instead, the evaluation of Kullback-Leibler divergence for an individual residual relies on the knowledge of the marginal density $p_i(e_i)$, where $e_i$ is the $i$-th residual for $i = 1, ..., n$. This is much easier to approximate through simulation. It is also worth mentioning that this independence assumption generally will not hold, since $\text{cov}(e_i, e_j) \neq 0$ if $\boldsymbol{R}_{ij} \neq 0$ for any $1 \leq i < j \leq n$, but its existence is essential for reducing the computational cost.

Given $\boldsymbol{X}$ and $\boldsymbol{\beta}$, the algorithm for approximating equation \ref{eq:kl-1} starts from simulating $m$ sets of observed values $\boldsymbol{y}$ according to the data generating process. The observed values are stored in a matrix $\boldsymbol{A}$ with $n$ rows and $m$ columns, where each column of $\boldsymbol{A}$ is a set of observed values. Then, we can get $m$ sets of realized values of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$. Furthermore, kernel density estimation (KDE) with Gaussian kernel and optimal bandwidth selected by the Silverman's rule of thumb [@silverman2018density] is applied on each row of $B$ to estimate $p_i(e_i)$ for $i = 1, ..., n$. The KDE computation can be done by the `density` function in R. 

Since the Kullback-Leibler divergence can be viewed as the expectation of the log-likelihood ratio between distribution $P$ and distribution $Q$ evaluated on distribution $P$, we can reuse the simulated residuals in matrix $\boldsymbol{B}$ to estimate the expectation by the sample mean. With the independence assumption, for non-normal $P$, $D_{KL}$ can be approximated by

\begin{align}
\label{eq:kl-3}
D_{KL} &\approx \sum_{i = 1}^{n} \hat{D}_{KL}^{(i)}, \\
\hat{D}_{KL}^{(i)} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
\end{align}

\noindent where $\hat{D}_{KL}^{(i)}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $B_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $B$, $\hat{p_i}(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\widehat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.

## Distance Estimation

In the previous sections, we have defined a distance measure given in equation \ref{eq:kl-0} for quantifying the difference between the true residual distribution $P$ and an ideal reference distribution $Q$. You may have noticed that this distance measure can only be computed when the data generating process is known. In reality, we often have no knowledge about the data generating process, otherwise we do not need to fit a linear regression model in the first place.

We tried to train a computer vision model to estimate this distance measure with a residual plot. Let $D$ be the result of equation \ref{eq:kl-0}, and our estimator $\hat{D}$ is formulated as 

\begin{equation}
\label{eq:d-approx}
\hat{D} = f_{CV}(V_{h \times w}(\boldsymbol{e}, \hat{\boldsymbol{y}})),
\end{equation}

\noindent where $V_{h \times w}(.)$ is a plotting function that saves a residuals vs fitted values plot with fixed aesthetic as an RGB image with $h \times w$ pixels, $f_{CV}(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance in the domain $[0, +\infty)$.

With the estimated distance $\hat{D}$, we will be able to know how different the underlying distribution of the residuals is from a good residual distribution. $\hat{D}$ can also be used as an index of the model violations indicating the strength of the visual signal embedded in the residual plot. 

It is not expected that $\hat{D}$ will be equal to original distance $D$. This is largely because information contained in a single residual plot is limited and it may not be able to summarise all the important characteristics of the residual distribution. For a given residual distribution $P$, many different residual plots can be simulated, where many will share similar visual patterns, but some of them could be visually very different from the rest, especially for regression models with small $n$. This suggests the error of the estimation will vary depends on whether the input residual plot is representative or not.

## Model Violations Index (MVI) {#sec:model-violations-index}

$\hat{D}$ is an estimator of the difference between the true residual distribution and the reference residual distribution. This difference primarily arises from deviations in model assumptions. The magnitude of $D$ directly reflects the degree of these deviations, thus making $\hat{D}$ instrumental in forming a model violations index.

Note that Equation \ref{eq:kl-0} might be influenced by the number of random variables involved in its evaluation. Generally, a larger number of observations will lead to a greater distance $D$. However, this does not imply that $\hat{D}$ fails to accurately represent the extent of model violations. In fact, when examining residual plots with more observations, we often observe a stronger visual signal strength, as the underlying patterns are more likely to be revealed, except in cases of significant overlapping. 

For a consistent data generating process, $D$ typically increases logarithmically with the number of observations. This behaviour comes from the relationship $D = \text{log}(1 + D_{KL})$, where $D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}$ under the assumption of independence.

Therefore, the Model Violations Index (MVI) can be proposed as

\begin{equation}
\label{eq:mvi}
\text{MVI} = C + \hat{D} - log(n),
\end{equation}

\noindent where $C$ is a large enough constant keeping the result positive. 

Figures \ref{fig:poly-index} and \ref{fig:heter-index} display the residual plots for fitted models exhibiting varying degrees of non-linearity and heteroskedasticity. Each residual plot's MVI is computed using Equation \ref{eq:mvi} with $C = 10$. When $\text{MVI} > 8$, the visual patterns are notably strong and easily discernible by humans. In the range $6 < \text{MVI} < 8$, the visibility of the visual pattern diminishes as MVI decreases. Conversely, when $\text{MVI} < 6$, the visual pattern tends to become relatively faint and challenging to observe. Table \ref{tab:mvi} provides a summary of the MVI usage and it is applicable to other linear regression models.

```{r mvi}
data.frame(degree = c("Strong", "Moderate", "Weak"),
           range = c("$\\text{MVI} > 8$", "$6 < \\text{MVI} < 8$", "$\\text{MVI} < 6$")) %>%
  kableExtra::kable("latex", 
                    booktabs = TRUE,
                    escape = FALSE,
                    col.names = c("Degree of model violations", "Range ($C$ = 10)"),
                    caption = "Degree of model violations or the strength of the visual signals according to the Model Violation Index (MVI). The constant $C$ is set to be 10.")
```




```{r poly-index, fig.pos = "!h", fig.cap = "Residual plots generated from fitted models exhibiting varying degrees of non-linearity violations. The Model violations index (MVI) is displayed atop each residual plot. The non-linearity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$."}
set.seed(1495)
if (!file.exists(here::here("paper/data/poly_index.rds"))) {
  e_sigma_factor <- 3
  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
  poly_dgp <- phn_model(j = 3, sigma = 0.1^e_sigma_factor)
  ori_dat <- poly_dgp$gen(300)
  
  vss <- c()
  actual_ss <- c()
  dat_combined <- tibble()
    
  for (e_sigma in seq(0.7, 1.8, length.out = 20)^e_sigma_factor) {
    poly_dgp <- phn_model(j = 3, sigma = e_sigma)
    dat <- poly_dgp$gen(300, computed = list(x1 = ori_dat$x1, e = ori_dat$e, k = ori_dat$k * e_sigma / 0.1^e_sigma_factor))
    mod <- lm(y ~ x1, data = dat)
    my_vi <- autovi::auto_vi(fitted_mod = mod, keras_mod = keras_mod)
    vss <- c(vss, my_vi$vss())
    actual_ss <- c(actual_ss, log(poly_dgp$sample_effect_size(dat) + 1))
    
    dat_combined <- bind_rows(dat_combined, mutate(dat, e_sigma = e_sigma))
  }
  
  saveRDS(dat_combined, here::here("paper/data/poly_index.rds"))
  saveRDS(vss, here::here("paper/data/poly_index_vss.rds"))
}

dat_combined <- readRDS(here::here("paper/data/poly_index.rds"))
vss <- readRDS(here::here("paper/data/poly_index_vss.rds"))

dat_combined %>%
  autovi::AUTO_VI$plot_resid() +
  facet_wrap(~e_sigma, scales = "free", 
             labeller = function(...){ 
               tibble(nb = map(glue("'MVI' == {format(vss - log(300) + 10, digits = 3)}"), str2expression))
               },
             ncol = 5) +
  theme(strip.text = element_text(size = 10))
```

```{r heter-index, fig.pos = "!h", fig.cap = "Residual plots generated from fitted models exhibiting varying degrees of heteroskedasticity violations. The Model violations index (MVI) is displayed atop each residual plot. The heteroskedasticity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$."}
set.seed(1495)
if (!file.exists(here::here("paper/data/heter_index.rds"))) {
  b_factor <- 2
  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
  heter_dgp <- phn_model(include_z = FALSE, b = 0)
  ori_dat <- poly_dgp$gen(300)
  
  vss <- c()
  actual_ss <- c()
  dat_combined <- tibble()
    
  for (b in rev(seq(0.2, 4, length.out = 20))^b_factor) {
    heter_dgp <- phn_model(include_z = FALSE, b = b)
    dat <- heter_dgp$gen(300, computed = list(x1 = ori_dat$x1, e = ori_dat$e))
    mod <- lm(y ~ x1, data = dat)
    my_vi <- autovi::auto_vi(fitted_mod = mod, keras_mod = keras_mod)
    vss <- c(vss, my_vi$vss())
    actual_ss <- c(actual_ss, log(heter_dgp$sample_effect_size(dat) + 1))
    
    dat_combined <- bind_rows(dat_combined, mutate(dat, nb = -b))
  }
  
  saveRDS(dat_combined, here::here("paper/data/heter_index.rds"))
  saveRDS(vss, here::here("paper/data/heter_index_vss.rds"))
}

dat_combined <- readRDS(here::here("paper/data/heter_index.rds"))
vss <- readRDS(here::here("paper/data/heter_index_vss.rds"))

dat_combined %>%
  autovi::AUTO_VI$plot_resid() +
  facet_wrap(~nb, scales = "free", 
             labeller = function(...){ 
               tibble(nb = map(glue("'MVI' == {format(vss - log(300) + 10, digits = 3)}"), str2expression))
               },
             ncol = 5) +
  theme(strip.text = element_text(size = 10))
```


## Statistical testing

### Lineup Evaluation

Theoretically, the distance $D$ for a correctly specified model is $0$, because $P$ will be the same as $Q$. However, the computer vision model may not necessary predict $0$ for a null plot. Using Figure \ref{fig:false-finding} as an example, it contains a visual pattern which is an indication of heteroskedasticity. We would not expect the model to be able to magically tell if the suspicious pattern is caused by the skewed distribution of the fitted values or the existence of heteroskedasticity. Additionally, some null plots could have outliers or strong visual patterns due to randomness, and a reasonable model will try to summarise those information into the prediction, resulting in $\hat{D} > 0$.

This property is not an issue if $\hat{D} \gg 0$ for which the visual signal of the residual plot is very strong, and we usually do not need any further examination of the significance of the result. However, if the visual pattern is weak or moderate, having $\hat{D}$ will not be sufficient to determine if $H_0$ should be rejected.

To address this issue while adhering to the principle of visual inference, we can compare the estimated distance $\hat{D}$ to the estimated distances for the null plots in a lineup. Specifically, if a lineup comprises 20 plots, the null hypothesis $H_0$ will be rejected if $\hat{D}$ exceeds the maximum estimated distance among the $m - 1$ null plots, denoted as $\max\limits_{1 \leq i \leq m-1} {\hat{D}_{null}^{(i)}}$, where $\hat{D}_{null}^{(i)}$ represents the estimated distance for the $i$-th null plot. This approach is equivalent to the typical lineup protocol requiring a 95% significance level, where $H_0$ is rejected if the data plot is identified as the most distinct plot by the sole participant. The estimated distance serves as a metric to quantify the difference between the data plot and the null plots, as intended.

For lineups consisting of more than 20 plots, the 95% significance level can be maintained if the number of plots is a multiple of 20. Specifically, for lineups comprising $20t$ plots, where $l$ is a positive integer, we reject $H_0$ if $\hat{D}$ exceeds 95% ${\hat{D}_{null}^{(i)}}$ for $i = 1, \ldots, 20t-1$. The $p$-value in this case is given by $\frac{1}{20t} + \frac{1}{20t}\sum{i=1}^{20t-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)$, where $I(\cdot)$ is the indicator function.

Moreover, if the number of plots in a lineup, denoted by $m$, is sufficiently large, the empirical distribution of ${\hat{D}_{null}^{(i)}}$ can be viewed as an approximation of the null distribution of the approximated distance. Consequently, quantiles of the null distribution can be estimated using the sample quantiles available in statistical software such as R, and these quantiles can be utilized for decision-making purposes. The details of the sample quantile computation can be found in @hyndman1996sample. For instance, if $\hat{D}$ is greater than or equal to the 95% sample quantile, denoted as $Q_{null}(0.95)$, we can conclude that the estimated distance for the true residual plot is significantly different from the estimated distance for null plots with a 95% significance level. Based on our experience, to obtain a stable estimate of the 95% quantile, the number of null plots, $n_{null}$, typically needs to be at least 100. However, if the null distribution exhibits a long tail, a larger number of null plots may be required. Alternatively, a $p$-value can be used to represent the probability of observing an event equally or more extreme than the given event under the null hypothesis $H_0$, and it can be estimated by $\frac{1}{m}\sum_{i=1}^{m-1}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)$.

If precision in sample quantiles is not the main priority, using a pre-calculated table of quantiles is an available option. Such a table offers pre-determined quantiles for a specified number of observations. It is generated by assessing numerous null plots derived from various simulated regression models and averaging them. Essentially, this shifts the computational burden from the user to the developer. 

### Bootstrapping

Bootstrap is often employed in linear regression when conducting inference for estimated parameters. It is typically done by sampling individual cases with replacement and refitting the regression model. If the observed data accurately reflects the true distribution of the population, the bootstrapped estimates can be used to measure the variability of the parameter estimate without making strong distributional assumptions about the data generating process.

Similarly, bootstrap can be applied on the estimated distance $\hat{D}$. For each refitted model $M_{boot}^{(i)}$, there will be an associated residual plot $V_{boot}^{(i)}$ which can be fed into the computer vision model to obtain $\hat{D}_{boot}^{(i)}$, where $i = 1,...,n_{boot}$, and $n_{boot}$ is the number of bootstrapped samples. If we are interested in the variation of $\hat{D}$, we can use $\hat{D}_{boot}^{(i)}$ to estimate a confidence interval. 

Alternatively, since each $M_{boot}^{(i)}$ has a set of estimated coefficients $\hat{\boldsymbol{\beta}}_{boot}^{(i)}$ and an estimated variance $\hat{\sigma^2}_{boot}^{(i)}$, a new approximated null distribution can be construed and the corresponding 95% sample quantile $Q_{boot}^{(i)}(0.95)$ can be computed. Then, if $\hat{D}_{boot}^{(i)} \geq Q_{boot}^{(i)}(0.95)$, $H_0$ will be rejected for $M_{boot}^{(i)}$. The ratio of rejected $M_{boot}^{(i)}$ among all the refitted models provides an indication of how often the assumed regression model are considered to be incorrect if the data can be obtained repetitively from the same data generating process. But this approach is computationally very expensive because it requires $n_{boot} \times n_{null}$ times of residual plot assessment. In practice, $Q_{null}(0.95)$ can be used to replace $Q_{boot}^{(i)}(0.95)$ in the computation. 

## Data Genearation

### Simulation Scheme

While observational data is frequently employed in training models for real-world applications, the data generating process of observational data often remains unknown, making computation for our target variable $D$ unattainable. Consequently, the computer vision models developed in this study were trained using synthetic data. This approach provided us with precise label annotations. Additionally, it ensured a large and diverse training dataset, as we had control over the data generating process, and the simulation of the training data was relatively cost-effective.

We have incorporated three types of residual departures of linear regression model in the training data, including non-linearity, heteroskedasticity and non-normality. All three departures can be summarised by the data generating process formulated as

\begin{align}
\label{eq:data-sim}
\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 + \beta_2(\boldsymbol{z} + \beta_1\boldsymbol{w}) + \boldsymbol{k} \odot \boldsymbol{\varepsilon}, \\
\boldsymbol{z} &= \text{He}_j(g(\boldsymbol{x}_1, 2)), \\
\boldsymbol{w} &= \text{He}_j(g(\boldsymbol{x}_2, 2)), \\
\boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2},
\end{align}

\noindent where $\boldsymbol{y}$, $\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{z}$, $\boldsymbol{w}$, $\boldsymbol{k}$ and $\boldsymbol{\varepsilon}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$ are two independent predictors, $\text{He}_j(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau], the $\sqrt{(.)}$ and $(.)^2$ operators are element-wise operators, $\odot$ is the Hadamard product, and $g(., k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as

$$g(\boldsymbol{x}, k) = 2k \cdot \frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k\boldsymbol{1}_n,~for~k > 0,$$
\noindent where $x_{min} = \underset{i \in \{ 1,...,n\}}{min} x_i$, $x_{max} = \underset{i \in \{ 1,...,n\}}{max} x_i$ and $x_i$ is the $i$-th entry of $\boldsymbol{x}$.

```{r factor}
tibble(Factor = c("j", "a", "b", "$\\beta_1$", "$\\beta_2$", " $\\text{dist}_\\varepsilon$", "$\\text{dist}_{x1}$", "$\\text{dist}_{x2}$", "$\\sigma_{\\varepsilon}$", "$\\sigma_{X1}$", "$\\sigma_{X2}$", "n"),
       Domain = c("\\{2, 3, ..., 18\\}", "[-1, 1]", "[0, 100]", "{0, 1}", "{0, 1}", "\\{discrete, uniform, normal, lognormal\\}", "\\{discrete, uniform, normal, lognormal\\}", "\\{discrete, uniform, normal, lognormal\\}", "[0.0625, 9]", "[0.3, 0.6]", "[0.3, 0.6]", "[50, 500]")) %>%
  kableExtra::kable("latex", 
                    escape = FALSE,
                    booktabs = TRUE,
                    caption = "Factors used in the data generating process for synthetic data simulation. Factor $j$ and $a$ controls the non-linearity shape and the heteroskedasticity shape respectively. Factor $b$, $\\sigma_\\varepsilon$ and $n$ control the signal strength. Factor $\\text{dist}_\\varepsilon$, $\\text{dist}_{x1}$ and $\\text{dist}_{x2}$ specifies the distribution of $\\varepsilon$, $X_1$ and $X_2$ respectively.")
```

The residuals and fitted values of the fitted model were obtained by regressing $\boldsymbol{y}$ on $\boldsymbol{x}_1$. If $\beta_1 \neq 0$, $\boldsymbol{x}_2$ was also included in the design matrix. This data generation process was adapted from @li2023plot, where it was utilized to simulate residual plots exhibiting non-linearity and heteroskedasticity visual patterns for human subject experiments. A summary of the factors utilized in Equation \ref{eq:data-sim} is provided in Table \ref{tab:factor}.

In Equation \ref{eq:data-sim}, $\boldsymbol{z}$ and $\boldsymbol{w}$ represent higher-order terms of $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$, respectively. If $\beta_2 \neq 0$, the regression model will encounter non-linearity issues. Parameter $j$ serves as a shape parameter that controls the number of tuning points in the non-linear pattern. Typically, higher values of $j$ lead to an increase in the number of tuning points, as illustrated in Figure \ref{fig:different-j}.

```{r different-j, fig.pos="!h", fig.cap = "Non-linearity forms generated for the synthetic data simulation. The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}

set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(j = 2, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
  mutate(j = 2)

# Generate data for shape 2, 3 and 4. Reuse x and e.
map_df(3:18, function(j) {
  phn_model(j = j, include_x2 = FALSE, sigma = 0.05)$
    gen(500, computed = select(dat_shape_1, x1, e)) %>%
  mutate(j = j)
}) %>%
  
  # Combined with data for shape 1
  bind_rows(dat_shape_1) %>%
  mutate(j = factor(j)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
```

Additionally, Scaling factor $\boldsymbol{k}$ directly affects the error distribution and it is correlated with $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$. If $b \neq 0$ and $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, the constant variance assumption will be violated. Parameter $a$ is a shape parameter controlling the location of the smallest variance in a residual plot as shown in Figure \ref{fig:different-a}.

```{r different-a, fig.width=8/5*3*2, fig.height=6/4*3*2, fig.pos="!h", fig.cap = 'Heteroskedasticity forms generated for the synthetic data simulation. Different shapes are controlled by the continuous factor $a$ between -1 and 1. For $a = -1$, the residual plot exhibits a "left-triangle" shape. And for $a = 1$, the residual plot exhibits a "right-triangle" shape. '}

set.seed(10085)

# Generate data for a = -1
dat_a_n1 <- phn_model(include_z = FALSE,
                      include_x2 = FALSE,
                      a = -1,
                      b = 100)$gen(500) %>%
  mutate(a = -1)

# Generate data for other a
map(c(-0.75, -0.5, -0.25, 0, 0.25, 0.5, 0.75, 1), function(a) {
  phn_model(include_z = FALSE,
            include_x2 = FALSE,
            a = a,
            b = 100)$gen(500) %>%
  mutate(a = a)
}) %>%
  
  # Combined with data for a = -1
  bind_rows(dat_a_n1) %>%
  mutate(a = factor(a)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~a, scales = "free", ncol = 3) +
  xlab("Fitted values") +
  ylab("Residuals")
```

```{r different-e, fig.pos="!h", fig.cap = 'Non-normality forms generated for the synthetic data simulation. Four different error distributions including discrete, lognormal, normal and uniform are considered.'}

set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_uniform(-1.4, 1.4))$gen(500) %>%
  mutate(e_dist = "uniform")

dat_shape_2 <- phn_model(include_z = FALSE, include_x2 = FALSE, sigma = 0.8)$gen(500) %>%
  mutate(e_dist = "normal")

dat_shape_3 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_lognormal(sigma = 0.6))$gen(500) %>%
  mutate(e_dist = "lognormal")

dat_shape_4 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_uniform_d(-1.4, 1.4, even = TRUE))$gen(500) %>%
  mutate(e_dist = "discrete")

# Generate data for shape 2, 3 and 4. Reuse x and e.
bind_rows(dat_shape_1, dat_shape_2, dat_shape_3, dat_shape_4) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~e_dist, scales = "free", labeller = label_parsed, ncol = 2)
```

Non-normality violations arise from specifying a non-normal distribution for $\boldsymbol{\varepsilon}$. In the synthetic data simulation, four distinct error distributions are considered, including discrete, uniform, normal, and log-normal distributions, as presented in Figure \ref{fig:different-e}. Each distribution imparts unique characteristics to the residuals. The discrete error distribution introduces discreteness in residuals, while the log-normal distribution typically yields outliers. Uniform error distribution may result in residuals filling the entire space of the residual plot. All of these distributions exhibit visual distinctions from the normal error distribution.

```{r different-j-x2, fig.pos="!h", fig.cap = "Residual plots of multiple linear regression models with non-linearity issues. The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$. A second predictor $\\boldsymbol{x}_2$ is introduced to the regression model to create complex shapes."}

set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(j = 2, include_x2 = TRUE, sigma = 0.05)$gen(500) %>%
  mutate(j = 2)

# Generate data for shape 2, 3 and 4. Reuse x and e.
map_df(3:18, function(j) {
  phn_model(j = j, include_x2 = TRUE, sigma = 0.05)$
    gen(500, computed = select(dat_shape_1, x1, e)) %>%
  mutate(j = j)
}) %>%
  
  # Combined with data for shape 1
  bind_rows(dat_shape_1) %>%
  mutate(j = factor(j)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
```

```{r different-j-heter, fig.pos="!h", fig.cap = 'Residual plots of models violating both the non-linearity and the heteroskedasticity assumptions. The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$, and the "left-triangle" shape is introduced by setting $a = -1$.'}

set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(j = 2, a = -1, b = 100, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
  mutate(j = 2)

# Generate data for shape 2, 3 and 4. Reuse x and e.
map_df(3:18, function(j) {
  phn_model(j = j, a = -1, b = 100, include_x2 = FALSE, sigma = 0.05)$
    gen(500, computed = select(dat_shape_1, x1, e)) %>%
  mutate(j = j)
}) %>%
  
  # Combined with data for shape 1
  bind_rows(dat_shape_1) %>%
  mutate(j = factor(j)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
```


```{r different-e-heter, fig.pos="!h", fig.cap = 'Residual plots of models violating both the non-normality and the heteroskedasticity assumptions. The four shapes are generated by using four different error distributions including discrete, lognormal, normal and uniform, and the "left-triangle" shape is introduced by setting $a = -1$. '}

set.seed(10085)

# Data for shape 1
dat_shape_1 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_uniform(-1.4, 1.4))$gen(500) %>%
  mutate(e_dist = "uniform")

dat_shape_2 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, sigma = 0.8)$gen(500) %>%
  mutate(e_dist = "normal")

dat_shape_3 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_lognormal(sigma = 0.6))$gen(500) %>%
  mutate(e_dist = "lognormal")

dat_shape_4 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_uniform_d(-1.4, 1.4, even = TRUE))$gen(500) %>%
  mutate(e_dist = "discrete")

# Generate data for shape 2, 3 and 4. Reuse x and e.
bind_rows(dat_shape_1, dat_shape_2, dat_shape_3, dat_shape_4) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~e_dist, scales = "free", labeller = label_parsed, ncol = 2)
```

Equation \ref{eq:data-sim} accommodates the incorporation of the second predictor $\boldsymbol{x}_2$. Introducing it into the data generation process by setting $\beta_1 = 1$ significantly enhances the complexity of the shapes, as illustrated in Figure \ref{fig:different-j-x2}. In comparison to Figure \ref{fig:different-j}, Figure \ref{fig:different-j-x2} demonstrates that the non-linear shape resembles a surface rather than a single curve. This augmentation can facilitate the computer vision model in learning visual patterns from residual plots of the multiple linear regression model.

In real-world analysis, it's not uncommon to encounter instances where multiple model violations coexist. In such cases, the residual plots often exhibit a mixed pattern of visual anomalies corresponding to different types of model violations. Figure \ref{fig:different-j-heter} and \ref{fig:different-e-heter} show the visual patterns of models with multiple model violations.

### Scagnostics


In Section \ref{introduction}, we mentioned that scagnostics consist of a set of manually designed visual feature extraction functions. While our computer vision model will learn its own feature extraction function during training, leveraging additional information from scagnostics can enhance the model's predictive accuracy.

For each generated residual plot, we computed four scagnostics  "Monotonic," "Sparse," "Splines," and "Striped"  using the `cassowaryr` R package [@mason2022cassowaryr]. These computed measures, along with the number of observations from the fitted model, were provided as the second input for the computer vision model. While other scagnostics provide valuable insights, they come with high computational costs and are not suitable for quick inference.

### Balanced Dataset

To train a robust computer vision model, we deliberately controlled the distribution of the target variable $D$ in the training data. We ensured that it followed a uniform distribution between $0$ and $7$. This was achieved by organizing $50$ buckets, each exclusively accepting training samples with $D$ falling within the range $[7(i - 1)/49, 7i/49)$ for $i < 50$, where $i$ represents the index of the $i$-th bucket. For the $50$-th bucket, any training samples with $D \geq 7$ were accepted.

With 80000 training images prepared, each bucket accommodated a maximum of $80000 \div 50 = 1600$ training samples. The simulator iteratively sampled parameter values from the parameter space, generated residuals and fitted values using the data generation process, computed the distance, and checked if the sample fitted within the corresponding bucket. This process continued until all buckets were filled.

Similarly, we adopted the same methodology to prepare 8000 test images for performance evaluation and model diagnostics.


## Model Architecture

```{r cnn-diag, out.width = "100%", fig.cap = "Diagram of the architecture of the optimized computer vision model."}
knitr::include_graphics("cnn.pdf")
```



The architecture of the computer vision model is adapted from a well-established architecture known as VGG16, which has demonstrated high performance in image classification [@simonyan2014very]. Figure \ref{fig:cnn-diag} provides a diagram of the architecture. More details about the neural network layers used in this study are provided in the Appendix.

The model begins with an input layer of shape $n \times h \times w \times 3$, capable of handling $n$ RGB images. This is followed by a grayscale conversion layer utilizing the luma formula under the Rec. 601 standard, which converts the colour image to grayscale. Grayscale suffices for our task since data points are plotted in black. We experiment with three combinations of $h$ and $w$: $32 \times 32$, $64 \times 64$, and $128 \times 128$, aiming to achieve sufficiently high image resolution for the problem at hand.

The processed image is used as the input for the first convolutional block. The model comprises at most five consecutive convolutional blocks, mirroring the original VGG16 architecture. Within each block, there are two 2D convolutional layers followed by two activation layers, respectively. Subsequently, a 2D max-pooling layer follows the second activation layer. The 2D convolutional layer convolves the input with a fixed number of $3 \times 3$ convolution filters, while the 2D max-pooling layer downsamples the input along its spatial dimensions by taking the maximum value over a $2 \times 2$ window for each channel of the input. The activation layer employs the rectified linear unit (ReLU) activation function, a standard practice in deep learning, which introduces a non-linear transformation of the output of the 2D convolutional layer. Additionally, to regularize training, a batch normalization layer is added after each 2D convolutional layer and before the activation layer. Finally, a dropout layer is appended at the end of each convolutional block to randomly set some inputs to zero during training, further aiding in regularization.

The output of the last convolutional block is summarized by either a global max pooling layer or a global average pooling layer, resulting in a two-dimensional tensor. To leverage the information contained in scagnostics, this tensor is concatenated with an additional $n \times 5$ tensor, which contains the "Monotonic," "Sparse," "Splines," and "Striped" measures, along with the number of observations for $n$ residual plots.

The concatenated tensor is then fed into the final prediction block. This block consists of two fully-connected layers. The first layer contains at least $128$ units, followed by a dropout layer. Occasionally, a batch normalization layer is inserted between the fully-connected layer and the dropout layer for regularization purposes. The second fully-connected layer consists of only one unit, serving as the output of the model.

The model weights $\boldsymbol{\theta}$ were randomly initialized and they were optimized by the Adam optimizer with the mean square error loss function

$$\hat{\boldsymbol{\theta}} = \underset{\boldsymbol{\theta}}{argmin}\frac{1}{n_{train}}\sum_{i=1}^{n_{train}}(D_i - f_{\boldsymbol{\theta}}(V_i, S_i))^2,$$

\noindent where $n_{train}$ is the number of training samples, $V_i$ is the $i$-th residual plot and $S_i$ is the additional information about the $i$-th residual plot including four scagnostics and the number of observations. 

## Model Training

To achieve an optimal deep learning model, hyperparameters like the learning rate often need to be fine-tuned using a tuner. In our study, we utilized the Bayesian optimization tuner from the `KerasTuner` Python library [@omalley2019kerastuner] for this purpose. A comprehensive list of hyperparameters is provided in Table \ref{tab:hyperparameter}.

The "number of base filters" determines the number of filters for the first 2D convolutional layer. In the VGG16 architecture, the number of filters for the 2D convolutional layer in a block is typically twice the number in the previous block, except for the last block, which maintains the same number of convolution filters as the previous one. This hyperparameter aids in controlling the complexity of the computer vision model. Higher numbers of base filters result in more trainable parameters. Likewise, the "number of units for the fully-connected layer" determines the complexity of the final prediction block. Increasing the number of units enhances model complexity, resulting in more trainable parameters.

The dropout rate and batch normalization are flexible hyperparameters that work in conjunction with other parameters to facilitate smooth training. A higher dropout rate is necessary when the model tends to overfit the training dataset by learning too much noise. Conversely, a lower dropout rate is preferred when the model is complex and challenging to converge. Batch normalization, on the other hand, addresses the internal covariate shift problem arising from the randomness in weight initialization and input data. It helps stabilize and accelerate the training process by normalizing the activations of each layer.

Additionally, incorporating additional inputs such as scagnostics and the number of observations can potentially enhance prediction accuracy. Therefore, we allowed the tuner to determine whether these inputs were necessary for optimal model performance.

The learning rate is a crucial hyperparameter, as it dictates the step size of the optimization algorithm. A high learning rate can help the model avoid local minima but risks overshooting and missing the global minimum. Conversely, a low learning rate smoothens the training process but makes the convergence time longer and increases the likelihood of getting trapped in local minima.

Our model was trained on the M3 high-performance computing platform, using TensorFlow and Keras. During training, 80% of the training data was utilized for actual training, while the remaining 20% was used as validation data. The Bayesian optimization tuner conducted 100 trials to identify the best hyperparameter values based on validation root mean square error. The tuner then restored the best epoch of the best model from the trials. Additionally, we applied early stopping, terminating the training process if the validation root mean square error fails to improve for 50 epochs. The maximum allowed epochs was set at 2000, although no models reached this threshold.


```{r hyperparameter}
data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
           Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[1e-8, 1e-1]")) %>%
  kableExtra::kable("latex", booktabs = TRUE, caption = "Name of hyperparameters and their correspoding domain for the computer vision model.")
```

Based on the tuning process described above, the optimized hyperparameter values are presented in Table \ref{tab:best-hyperparameter}. It was observable that a minimum of $32$ base filters was necessary, with the preferable choice being $64$ base filters for both the $64 \times 64$ and $128 \times 128$ models, mirroring the original VGG16 architecture. The optimized dropout rate for convolutional blocks hovered around $0.4$, and incorporating batch normalization for convolutional blocks proved beneficial for performance.

All optimized models chose to retain the additional inputs, contributing to the reduction of validation error. The number of units required for the fully-connected layer was $256$, a relatively modest number compared to the VGG16 classifier, suggesting that the problem at hand was less complex. The optimized learning rates were smaller than the default value recommended by Keras, which is 0.001.


```{r best-hyperparameter}
data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
           `32` = c("32", "0.4", "true", "max", "false", "256", "false", "0.2", "0.0003"),
           `64` = c("64", "0.3", "true", "average", "false", "256", "true", "0.4", "0.0006"),
           `128` = c("64", "0.4", "true", "average", "false", "256", "true", "0.1", "0.0052")) %>%
  kableExtra::kable("latex", 
                    booktabs = TRUE, 
                    caption = "Hyperparameters values for the optimized computer vision models with different input sizes.", 
                    col.names = c("Hyperparameter", "$32 \\times 32$", "$64 \\times 64$", "$128 \\times 128$"),
                    escape = FALSE) %>%
  kableExtra::kable_styling(latex_options = c("scale_down"))
```

<!-- ## Evaluation Methods -->

<!-- This section remains incomplete as we may modify the "Results" section later on. -->

# Results

## Model Performance

The training and test performance for the optimized models with three different input sizes are summarized in Table \ref{tab:performance}. Among these models, the $64 \times 64$ model and the $32 \times 32$ model consistently exhibited the best training and test performance, respectively. The mean absolute error indicated that the difference between $\hat{D}$ and $D$ was approximately $0.43$ on the test set, a negligible deviation considering the normal range of $D$ typically falls between $0$ and $7$. The high $R^2$ values also suggested that the predictions were largely linearly correlated with the target.

Figure \ref{fig:model-performance} presents a hexagonal heatmap for $D - \hat{D}$ versus $\hat{D}$. The red smoothing curves, fitted by generalized additive models [@hastie2017generalized], demonstrate that all the optimized models perform admirably on both the training and test sets. No structural issues are noticeable in Figure \ref{fig:model-performance}, but some minor issues regarding over-prediction and under-prediction are observed.

The figure highlights that most under-predictions occurred when $\hat{D} < 3$, while over-predictions occurred predominantly when $3 < \hat{D} < 6$. For input images representing null plots, it was expected that the models will over-predict the distance, as explained in Section \ref{lineup-evaluation}. However, it can not explain the under-prediction issue. Therefore, we analysed the relationship between residuals and all the factors involved in the data generating process. We found that most issues actually arose from non-linearity problems and the presence of a second predictor in the regression model as illustrated in Figure \ref{fig:over-under}. When the variance for the error distribution was small, the optimized model tended to under-predict the distance. Conversely, when the error distribution had a large variance, the model tended to over-predict the distance.  

Since most of the deviation stemmed from the presence of non-linearity violations, to further investigate this, we split the test set based on violation types and re-evaluated the performance, as detailed in Table \ref{tab:performance-sub}. It was evident that metrics for null plots were notably worse compared to other categories. Furthermore, residual plots solely exhibiting non-normality issues were the easiest to predict, with very low test root mean square error (RMSE) at around $0.3$. Residual plots with non-linearity issues were more challenging to assess than those with heteroskedasticity or non-normality issues. Assessing residual plots with heteroskedasticity was not as difficult as assessing those with non-linearity issues. When multiple violations were introduced to a residual plot, the performance metrics typically lay between the metrics for each individual violation.

Based on the model performance metrics, we chose to use the best-performing model evaluated on the test set, namely the $32 \times 32$ model, for the subsequent analysis.


```{r performance}
test_pred <- readr::read_csv(here::here("paper/data/test_pred.csv"))
train_pred <- readr::read_csv(here::here("paper/data/train_pred.csv"))
meta <- readr::read_csv(here::here("paper/data/meta.csv"))


test_summary <- test_pred %>% 
  left_join(meta) %>%
  group_by(res) %>%
  summarise(RMSE = yardstick::rmse_vec(effect_size, vss),
            R2 = yardstick::rsq_vec(effect_size, vss),
            MAE = yardstick::mae_vec(effect_size, vss),
            HUBER = yardstick::huber_loss_vec(effect_size, vss)) %>%
  mutate(type = "test")

train_summary <- train_pred %>% 
  left_join(meta) %>%
  group_by(res) %>%
  summarise(RMSE = yardstick::rmse_vec(effect_size, vss),
            R2 = yardstick::rsq_vec(effect_size, vss),
            MAE = yardstick::mae_vec(effect_size, vss),
            HUBER = yardstick::huber_loss_vec(effect_size, vss)) %>%
  mutate(type = "train")

train_summary %>%
  bind_rows(test_summary) %>%
  select(res, RMSE, R2, MAE, HUBER) %>%
  mutate(res = glue::glue("${res} \\times {res}$")) %>%
  mutate(across(RMSE:MAE, ~format(.x, digits = 3))) %>%
  mutate(across(HUBER, ~format(.x, digits = 2))) %>%
  kableExtra::kable(format = "latex", 
                    booktabs = TRUE, 
                    escape = FALSE,
                    col.names = c("", "RMSE", "$R^2$", "MAE", "Huber loss"),
                    caption = "The training and test performance of three optimized models with different input sizes.") %>%
  kableExtra::pack_rows("Training set", 1, 3) %>%
  kableExtra::pack_rows("Test set", 4, 6)
```


```{r model-performance, fig.pos="!h", fig.cap = "Hexagonal heatmap for residuals vs predicted distance on training and test data for three optimized models with different input sizes. The red lines are smoothing curves produced by fitting gnealized additive models. The area over the zero line in light pink indicates under-prediction, and the area under the zero line in light green indicates over-prediction.", dev = 'png', dpi = 300}
model_pred <- train_pred %>% 
  left_join(meta) %>%
  mutate(type = "train") %>%
  bind_rows(test_pred %>% 
              left_join(meta) %>%
              mutate(type = "test"))

model_pred %>%
  ggplot() +
  geom_hline(yintercept = 0, alpha = 0.5) +
  # geom_point(aes(effect_size, effect_size - vss), alpha = 0.03) +
  annotate("rect", 
           ymin = min(model_pred$effect_size - model_pred$vss),
                ymax = 0,
                xmin = min(model_pred$vss),
                xmax = max(model_pred$vss),
            fill = "#66c2a5",
            alpha = 0.3) +
  geom_point(data = NULL, aes(5, 0, col = "over-prediction"), shape = 15) +
  annotate("rect", 
           ymin = 0,
                ymax = max(model_pred$effect_size - model_pred$vss),
                xmin = min(model_pred$vss),
                xmax = max(model_pred$vss),
            fill = "#fc8d62",
            alpha = 0.3) +
  geom_point(data = NULL, aes(5, 0, col = "under-prediction"), shape = 15) +
  geom_hex(aes(vss, effect_size - vss), bins = 25) +
  geom_smooth(aes(vss, effect_size - vss), se = FALSE, col = "#e34a33") +
  facet_grid(type ~ res) +
  coord_fixed() +
  ylab(expression(D - hat(D))) +
  xlab(expression(hat(D))) +
  theme_light() +
  scale_fill_continuous(limits = c(1, 200), trans = "log10", low = "#56B1F7", high = "#132B43") +
  scale_color_manual(values = c("under-prediction" = scales::alpha("#fc8d62", 0.3),
                                "over-prediction" = scales::alpha("#66c2a5", 0.3)),
                     breaks = c("under-prediction", "over-prediction")) +
  guides(col = guide_legend(title = NULL, override.aes = list(size = 8), order = 1))
```


```{r over-under, fig.pos="!h", fig.cap = "Scatter plots for residuals vs $\\sigma$ on test data for the $32 \\times 32$ optimized model. The data is grouped by whether the regression has only non-linearity violation, and whether it includes a second predictor in the regression formula. The red lines are smoothing curves produced by fitting gnealized additive models. The area over the zero line in light pink indicates under-prediction, and the area under the zero line in light green indicates over-prediction.", dev = 'png', dpi = 300}
model_pred %>%
  filter(res == 32L) %>%
  filter(type == "test") %>%
  mutate(only_z = include_z & !include_heter & !include_non_normal) %>%
  rename(`Only has non-linearity violation` = only_z, `Has second predictor` = include_x2) %>%
  ggplot() +
  geom_hline(yintercept = 0, alpha = 0.5) +
  annotate("rect", 
           ymin = min(model_pred$effect_size[model_pred$type == "test"] - model_pred$vss[model_pred$type == "test"]),
                ymax = 0,
                xmin = min(model_pred$e_sigma),
                xmax = max(model_pred$e_sigma),
            fill = "#66c2a5",
            alpha = 0.3) +
  annotate("rect", 
           ymin = 0,
                ymax = max(model_pred$effect_size[model_pred$type == "test"] - model_pred$vss[model_pred$type == "test"]),
                xmin = min(model_pred$e_sigma),
                xmax = max(model_pred$e_sigma),
            fill = "#fc8d62",
            alpha = 0.3) +
  geom_point(aes(e_sigma, effect_size - vss), alpha = 0.2) +
  geom_smooth(aes(e_sigma, effect_size - vss), se = FALSE, col = "#e34a33") +
  facet_grid(`Has second predictor` ~ `Only has non-linearity violation`,
             labeller = label_both) +
  theme_light() +
  ylab(expression(D - hat(D))) +
  xlab(expression(sigma))
```


```{r performance-sub}

model_pred %>%
  filter(res == 32L) %>%
  group_by(include_non_normal, include_heter, include_z) %>%
  summarise(train_n = sum(type == "train"),
            train_RMSE = yardstick::rmse_vec(effect_size[type == "train"], vss[type == "train"]),
            test_n = sum(type == "test"),
            test_RMSE = yardstick::rmse_vec(effect_size[type == "test"], vss[type == "test"])) %>%
  mutate(violations = ifelse(include_z, "non-linearity", "null")) %>%
  mutate(violations = ifelse(include_heter, glue("{violations} + heteroskedasticity"), violations)) %>%
  mutate(violations = ifelse(include_non_normal, glue("{violations} + non-normality"), violations)) %>%
  mutate(violations = gsub("null \\+ ", "", violations)) %>%
  ungroup() %>%
  select(violations, train_n, train_RMSE, test_n, test_RMSE) %>%
  mutate(train_n = format(train_n)) %>%
  mutate(test_n = format(test_n)) %>%
  mutate(across(c(train_RMSE, test_RMSE), ~format(.x, digits = 3))) %>%
  mutate(violations = ifelse(violations == "null", "no violations", violations)) %>%
  kableExtra::kable("latex", 
                    booktabs = TRUE,
                    escape = FALSE,
                    col.names = c("Violations", "\\#training samples", "Training RMSE", "\\#test samples", "Test RMSE"),
                    caption = "The training and test performance of the $32 \\times 32$ model presented with different model violations.") %>%
  kableExtra::kable_styling(latex_options = c("scale_down"))
```


<!-- - check if j has an impact on residuals more carefully -->
<!-- - the RMSE has very little to do with j if x2 is not included. And its impact is not as large as sigma. -->


## Comparison with Human Visual Inference and Conventional Tests

### Overview of the Human Subject experiment

In order to check the validity of the proposed computer vision model, residual plots presented in the human subject experiment conducted by @li2023plot will be assessed.


<!-- The experiment revealed that conventional tests are more sensitive to weak departures from model assumptions than visual test, and they often reject the null hypothesis when departures are not visibly different from null residual plots. -->

This study has collected $7974$ human responses to $1152$ lineups. Each lineup contains one randomly placed true residual plot and 19 null plots. Among the $1152$ lineups, $24$ are attention check lineups in which the visual patterns are designed to be extremely obvious and very different from the corresponding to null plots, $36$ are null lineups where all the lineups consist of only null plots, $279$ are lineups with uniform predictor distribution evaluated by $11$ participants, and the remaining $813$ are lineups with discrete, skewed or normal predictor distribution evaluated by $5$ participants. Attention check lineups and null lineups will not be assessed in the following analysis. 

In @li2023plot, the residual plots are simulated from a data generating process which is a special case of Equation \ref{eq:data-sim}. The main characteristic is the model violations are introduced separately, meaning non-linearity and heteroskedasticity will not co-exist in one lineup but assigned uniformly to all lineups. Additionally, non-normality and multiple predictors are not considered in the experimental design.


### Model performance on the human data

```{r rd-human}
vss_32 <- readRDS(here::here("paper/data/vss_32.rds"))

experiment <- vi_survey %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  select(unique_lineup_id, attention_check, null_lineup, prop_detect,
         answer, effect_size, conventional_p_value,
         p_value, type, shape, a, b, x_dist, e_dist,
         e_sigma, include_z, k, n)

experiment <- vss_32 %>%
  left_join(experiment) %>%
  mutate(conventional_reject = conventional_p_value <= 0.05) %>%
  mutate(reject = p_value <= 0.05) %>%
  mutate(model_reject = vss_p_value <= 0.05) %>%
  mutate(model_boot_reject = vss_boot_p_value <= 0.05)

experiment <- experiment %>%
  left_join(readRDS(here::here("paper/data/actual_ss.rds")))
```

```{r experiment-performance}
experiment %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  group_by(type) %>%
  summarise(RMSE = yardstick::rmse_vec(actual_ss, vss),
            R2 = yardstick::rsq_vec(actual_ss, vss),
            MAE = yardstick::mae_vec(actual_ss, vss),
            HUBER = yardstick::huber_loss_vec(actual_ss, vss)) %>%
  mutate(across(RMSE:HUBER, ~format(.x, digits = 3))) %>%
  kableExtra::kable(format = "latex", 
                    booktabs = TRUE, 
                    escape = FALSE,
                    col.names = c("Voilation", "RMSE", "$R^2$", "MAE", "Huber loss"),
                    caption = "The performance of the $32 \\times 32$ model on the data used in the human subject experiment.")
```

For each lineup used in @li2023plot, there is one true residual plot and 19 null plots. While the distance $D$ for the true residual plot depends on the underlying data generating process, the distance $D$ for the null plots is zero. We have used our optimized computer vision model to estimate distance for both the true residual plots and the null plots. To have a fair comparison, $H_0$ will be rejected if the true residual plot has the greatest approximated distance among all plots in a lineup. Additionally, the appropriate conventional tests including the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey1969tests] for non-linearity and the Breusch-Pagan test  [@breusch1979simple] for heteroskedasticity were applied on the same data for comparison.

The performance metrics of $\hat{D}$ for true residual plots are outlined in Table \ref{tab:experiment-performance}. It's notable that all performance metrics are slightly worse than those evaluated on the test data. Nevertheless, the mean absolute error remains at a low level, and the linear correlation between the prediction and the true value remains very high. Lineups with non-linearity issues are more challenging to predict than those with heteroskedasticity issues.

<!-- For null plots, a histogram of the predictions is provided in Figure \ref{fig:hist-null-human}. In most cases, $\hat{D}$ is predicted to be around $1$. As $\hat{D}$ increases, the number of predictions decreases. The 90% and 95% sample quantiles are approximately at $2$, and the 99% sample quantile is around $2.5$. These sample quantiles can serve as global critical values for determining whether a residual plot needs to be rejected. We have also computed the 95% sample quantiles by assessing 200 null plots for each fitted model used in the experiment. The difference between decisions based on these two types of critical values is minor, with decisions aligning `r experiment %>% filter(!attention_check) %>% filter(!null_lineup) %>% summarise(q = mean(model_reject == (2.03 <= vss)) * 100) %>% pull(q) %>% format(digits = 3)`% of the time. Therefore, we will be using the 95% sample quantiles obtained from each fitted model in the following analysis.  -->

Table \ref{tab:human-conv-table} provides a summary of the agreement between decisions made by the computer vision model and conventional tests. The agreement rates between conventional tests and the computer vision model are 85.95% and 79.69% for residual plots containing heteroskedasticity and non-linearity patterns, respectively. These figures are higher than those calculated for visual tests conducted by human, indicating that the computer vision model exhibits behaviour more akin to the best available conventional tests. However, Figure \ref{fig:conv-mosaic} shows that the computer vision model does not always reject when the conventional tests reject. And a small number of plots will be rejected by computer vision model but not by conventional tests.  This suggests that conventional tests are more sensitive than the computer vision model.

Figure \ref{fig:pcp} further illustrates the decisions made by visual tests conducted by human, computer vision models, and conventional tests, using a parallel coordinate plots. It can be observed that all three tests will agree with each other for around 50% of the cases. When visual tests conducted by human do not reject, there are substantial amount of cases where computer vision model also not reject but conventional tests reject. There are much fewer cases that do not reject by visual tests and conventional tests, but reject by computer vision models. This indicates computer vision model can behave like visual tests conducted by human better than conventional tests. Moreover, there are great proportion of cases where visual tests conducted by human is the only test who does not reject, . 


When plotting the decision against the distance, as illustrated in Figure \ref{fig:power}, several observations emerge. Firstly, for the computer vision model, when the distance $D > 4$, almost all residual plots are rejected. However, for visual tests conducted by humans, the threshold is higher, set at $D > 5$. And conventional tests exhibit some non-rejection cases for $D > 3$. Moreover, the computer vision model tends to have fewer rejected cases than conventional tests when $D < 2$. However, visual tests demonstrate the lowest sensitivity to residual plots with small distances. Therefore, tests based on the computer vision model are less sensitive to small deviations from model assumptions than conventional tests but more sensitive to moderate deviations. Rejection decisions are fitted by logistic regression models with no intercept terms but an offset equals to $\text{log}(0.05/0/95)$. The fitting curves for the computer vision model fall between those of conventional tests and visual tests for both non-linearity and heteroskedasticity. There is still potential to refine the computer vision model to better align its behaviour with visual tests.



Figure \ref{fig:delta} displays the scatter plot of the weighted detection rate vs the $\delta$-difference. In the experiment conducted in @li2023plot, participants were allowed to make multiple selections for a lineup. The weighted detection rate is computed by assigning weights to each detection. If the participant selects zero plots, the corresponding weight is 0.05; otherwise, the weight is 1 divided by the number of selections. The $\delta$-difference is defined by @chowdhury2018measuring as

\begin{equation}
\delta = \hat{D} - \underset{j}{max}\left(\hat{D}_{null}^{(j)}\right) \quad \text{for}~j = 1,...,m-1,
\end{equation}

\noindent where $\hat{D}_{null}^{(j)}$ is the approximated distance for the $j$-th null plot, and $m$ is the number of plots in a lineup.

Figure \ref{fig:delta} indicates that the weighted detection rate increases as the $\delta$-difference increases, particularly when the $\delta$-difference is greater than zero. A negative $\delta$-difference suggests that there is at least one null plot in the lineup with a stronger visual signal than the true residual plot. In some instances, the weighted detection rate is close to one, yet the $\delta$-difference is negative. This discrepancy implies that the distance measure, or the approximated distance, may not perfectly reflect actual human behaviour.

```{r hist-null-human, fig.pos =  "!h", fig.cap = "Histogram of predicted distance on null plots used in the human subject experiment. Sample quantiles are drawn in different colors."}
lineup_vss <- readRDS(here::here("paper/data/lineup_vss.rds"))
lineup_vss <- lineup_vss %>%
  group_by(unique_lineup_id) %>%
  mutate(delta_diff = vss[!null] - max(vss[null])) %>%
  mutate(gamma_diff = sum(vss[null] > vss[!null]))


experiment <- experiment %>%
  left_join(lineup_vss %>%
  filter(null == FALSE) %>%
  mutate(model_reject_20 = rank == 1) %>%
  select(unique_lineup_id, rank, model_reject_20))

# lineup_vss %>%
#   filter(null) %>%
#   left_join(vi_survey %>%
#               group_by(unique_lineup_id) %>% 
#               summarise(type = first(type), 
#                         attention_check = first(attention_check),
#                         null_lineup = first(null_lineup))) %>%
#   filter(!attention_check) %>%
#   filter(!null_lineup) %>%
#   mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
#   ggplot() +
#   geom_histogram(aes(vss)) +
#   geom_vline(aes(xintercept = quantile(vss, 0.9), col = "Q(0.90)")) +
#   geom_vline(aes(xintercept = quantile(vss, 0.95), col = "Q(0.95)")) +
#   geom_vline(aes(xintercept = quantile(vss, 0.99), col = "Q(0.99)")) +
#   xlab(expression(hat(D))) +
#   theme_light() +
#   facet_wrap(~type) +
#   labs(col = "")
```

```{r cache = TRUE}
vi_lineup <- get_vi_lineup()
```


```{r eval = FALSE}
library(visage)
conv_experiment <- vi_lineup %>%
  map_df(function(.x) {
    reset_p_value <- c()
    bp_p_value <- c()
    sw_p_value <- c()
    
    for (i in 1:20) {
      this_dat <- .x$data %>% filter(k == i)
      reset_p_value <- c(reset_p_value, POLY_MODEL$test(this_dat, test = "RESET")$p_value)
      bp_p_value <- c(bp_p_value, HETER_MODEL$test(this_dat)$p_value)
      sw_p_value <- c(sw_p_value, shapiro.test(this_dat$.resid)$p.value)
    }
    return(tibble(unique_lineup_id = rep(.x$metadata$name, 20),
                  ans = .x$metadata$answer,
                  pos = 1:20,
                  reset_p_value = reset_p_value,
                  reset_rank = rank(reset_p_value, ties.method = "first"),
                  bp_p_value = bp_p_value,
                  bp_rank = rank(bp_p_value, ties.method = "first"),
                  sw_p_value = sw_p_value,
                  sw_rank = rank(sw_p_value, ties.method = "first")))
  })

conv_experiment %>%
  filter(ans == pos) %>%
  mutate(reset_reject_20 = reset_rank == 1,
         bp_reject_20 = bp_rank == 1,
         sw_reject_20 = sw_rank == 1) %>%
  select(-ans, -pos, -c(reset_p_value:sw_rank)) %>%
  right_join(experiment) %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  pivot_longer(c(reject, model_reject_20, reset_reject_20, bp_reject_20, conventional_reject, model_reject)) %>%
  select(type, actual_ss, name, value) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  nest(dat = c(offset0, actual_ss, value)) %>%
  mutate(mod = map(dat, 
                   ~glm(value ~ actual_ss - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(actual_ss = seq(0, 10, 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  ggplot() +
  geom_line(aes(actual_ss, power, col = name)) +
  facet_wrap(~type, scales = "free_x")
```


```{r human-conv-table, eval = TRUE}
bind_rows(
  experiment %>%
    filter(!attention_check) %>%
    filter(!null_lineup) %>%
    mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
    group_by(type) %>%
    summarise(n = format(n()), agree = format(sum(model_reject == conventional_reject)), rate = format(mean(conventional_reject == model_reject), digits = 4)),
  experiment %>%
    filter(!attention_check) %>%
    filter(!null_lineup) %>%
    mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
    group_by(type) %>%
    summarise(n = format(n()), agree = format(sum(model_reject == reject)), rate = format(mean(model_reject == reject), digits = 4))
  ) %>%
  kableExtra::kable("latex", 
                    booktabs = TRUE,
                    escape = FALSE,
                    col.names = c("Violations", "\\#Samples", "\\#Agreements", "Agreement rate"),
                    caption = "Summary of the comparison of decisions made by computer vision model with decisions made by conventional tests and visual tests conducted by human.") %>%
  kableExtra::pack_rows("Compared with conventional tests", 1, 2) %>%
  kableExtra::pack_rows("Compared with visual tests conducted by human", 3, 4)
```


```{r human-mosaic, fig.pos = "!h", fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of visual test conditional on the computer vision model decision on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot. The visual test rejects less frequently than the computer vision model, and (almost) only rejects when the computer vision model does.", eval = FALSE}
library(ggmosaic)

experiment %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  mutate(type = ifelse(type == "non-linearity", "Non-linearity", "Heteroskedasticity")) %>%
  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
  mutate(model_reject_20 = ifelse(model_reject_20, "Reject", "No")) %>%
  mutate(reject = ifelse(reject, "Reject", "No")) %>%
  mutate(across(c(model_reject_20, reject), ~factor(.x, levels = c("Reject", "No")))) %>%
  ggplot() +
  geom_mosaic(aes(x = ggmosaic::product(reject, model_reject_20), 
                  fill = model_reject_20)) +
  facet_grid(~type) +
  ylab("Visual tests reject") +
  xlab("Computer vision model rejects") +
  # labs(fill = "Visual tests reject") +
  scale_fill_brewer("", palette = "Dark2") +
  theme_bw() +
  theme(legend.position = "none") +
  coord_fixed()
```

```{r conv-mosaic, fig.pos = "!h", fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of computer vision models conditional on conventional tests on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot."}
library(ggmosaic)

experiment %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  mutate(type = ifelse(type == "non-linearity", "Non-linearity", "Heteroskedasticity")) %>%
  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
  mutate(model_reject_20 = ifelse(model_reject_20, "Reject", "No")) %>%
  mutate(conventional_reject = ifelse(conventional_reject, "Reject", "No")) %>%
  mutate(across(c(model_reject_20, conventional_reject), ~factor(.x, levels = c("Reject", "No")))) %>%
  ggplot() +
  geom_mosaic(aes(x = ggmosaic::product(model_reject_20, conventional_reject), 
                  fill = model_reject_20)) +
  facet_grid(~type) +
  ylab("Computer vision model rejects") +
  xlab("Conventional tests reject ") +
  # labs(fill = "Visual tests reject") +
  scale_fill_brewer("", palette = "Dark2") +
  theme_bw() +
  theme(legend.position = "none") +
  coord_fixed()
```


```{r power, fig.pos = "!h", fig.cap = "Comparison of power of visual tests, conventional tests and the computer vision model. Marks along the x-axis at the bottom of the plot represent rejections made by each type of test. Marks at the top of the plot represent acceptances. Power curves are fitted by logistic regression models with no intercept but an offset equals to $\\text{log}(0.05/0.95)$."}

max_actual_ss <- max(filter(experiment, !attention_check, !null_lineup)$actual_ss)
max_poly_ss <- max(filter(experiment, !attention_check, !null_lineup, type == "polynomial")$actual_ss)
max_heter_ss <- max(filter(experiment, !attention_check, !null_lineup, type != "polynomial")$actual_ss)

model_glm_pred_poly <- glm(model_reject_20 ~ actual_ss - 1, 
    data = filter(experiment, !attention_check, !null_lineup, type == "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
    offset = offset0,
    family = binomial()) %>%
  predict(data.frame(actual_ss = seq(0, max_poly_ss, 0.01), offset0 = log(0.05/0.95)),
          type = "response") %>%
  data.frame(d = seq(0, max_poly_ss, 0.01), type = "Non-linearity", name = "Computer vision model", value = .) %>%
  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))

visual_glm_pred_poly <- glm(reject ~ actual_ss - 1, 
    data = filter(experiment, !attention_check, !null_lineup, type == "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
    offset = offset0,
    family = binomial()) %>%
  predict(data.frame(actual_ss = seq(0, max_poly_ss, 0.01), offset0 = log(0.05/0.95)),
          type = "response") %>%
  data.frame(d = seq(0, max_poly_ss, 0.01), type = "Non-linearity", name = "Visual test", value = .) %>%
  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))

conv_glm_pred_poly <- glm(conventional_reject ~ actual_ss - 1, 
    data = filter(experiment, !attention_check, !null_lineup, type == "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
    offset = offset0,
    family = binomial()) %>%
  predict(data.frame(actual_ss = seq(0, max_poly_ss, 0.01), offset0 = log(0.05/0.95)),
          type = "response") %>%
  data.frame(d = seq(0, max_poly_ss, 0.01), type = "Non-linearity", name = "Conventional test", value = .) %>%
  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))

model_glm_pred_heter <- glm(model_reject_20 ~ actual_ss - 1, 
    data = filter(experiment, !attention_check, !null_lineup, type != "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
    offset = offset0,
    family = binomial()) %>%
  predict(data.frame(actual_ss = seq(0, max_heter_ss, 0.01), offset0 = log(0.05/0.95)),
          type = "response") %>%
  data.frame(d = seq(0, max_heter_ss, 0.01), type = "Heteroskedasticity", name = "Computer vision model", value = .) %>%
  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))

visual_glm_pred_heter <- glm(reject ~ actual_ss - 1, 
    data = filter(experiment, !attention_check, !null_lineup, type != "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
    offset = offset0,
    family = binomial()) %>%
  predict(data.frame(actual_ss = seq(0, max_heter_ss, 0.01), offset0 = log(0.05/0.95)),
          type = "response") %>%
  data.frame(d = seq(0, max_heter_ss, 0.01), type = "Heteroskedasticity", name = "Visual test", value = .) %>%
  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))

conv_glm_pred_heter <- glm(conventional_reject ~ actual_ss - 1, 
    data = filter(experiment, !attention_check, !null_lineup, type != "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
    offset = offset0,
    family = binomial()) %>%
  predict(data.frame(actual_ss = seq(0, max_heter_ss, 0.01), offset0 = log(0.05/0.95)),
          type = "response") %>%
  data.frame(d = seq(0, max_heter_ss, 0.01), type = "Heteroskedasticity", name = "Conventional test", value = .) %>%
  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))

mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
  filter(!null_lineup, !attention_check) %>%
ggplot() +
  geom_line(data = model_glm_pred_poly, aes(d, value, col = "Computer vision model"), linewidth = 0.8) +
  geom_line(data = model_glm_pred_heter, aes(d, value, col = "Computer vision model"), linewidth = 0.8) +
  geom_line(data = visual_glm_pred_poly, aes(d, value, col = "Visual test"), linewidth = 0.8) +
  geom_line(data = visual_glm_pred_heter, aes(d, value, col = "Visual test"), linewidth = 0.8) +
  geom_line(data = conv_glm_pred_poly, aes(d, value, col = "Conventional test"), linewidth = 0.8) +
  geom_line(data = conv_glm_pred_heter, aes(d, value, col = "Conventional test"), linewidth = 0.8) +
  geom_segment(data = ~filter(.x, !reject), aes(x = actual_ss, xend = actual_ss, y = -0.1, yend = -0.1 + 0.03, col = "Visual test"), alpha = 0.4) +
  geom_segment(data = ~filter(.x, reject), aes(x = actual_ss, xend = actual_ss, y = 1.1 - 0.03 * 2, yend = 1.1 - 0.03 * 3, col = "Visual test"), alpha = 0.4) +
  geom_segment(data = ~filter(.x, !model_reject_20), aes(x = actual_ss, xend = actual_ss, y = -0.1 + 0.03, yend = -0.1 + 0.03 * 2, col = "Computer vision model"), alpha = 0.4) +
  geom_segment(data = ~filter(.x, model_reject_20), aes(x = actual_ss, xend = actual_ss, y = 1.1 - 0.03, yend = 1.1 - 0.03 * 2, col = "Computer vision model"), alpha = 0.4) +
  geom_segment(data = ~filter(.x, !conventional_reject), aes(x = actual_ss, xend = actual_ss, y = -0.1 + 0.03 * 2, yend = -0.1 + 0.03 * 3, col = "Conventional test"), alpha = 0.4) +
  geom_segment(data = ~filter(.x, conventional_reject), aes(x = actual_ss, xend = actual_ss, y = 1.1, yend = 1.1 - 0.03, col = "Conventional test"), alpha = 0.4) +
  facet_grid(~type, scales = "free_x") +
  theme_light() +
  scale_color_brewer("", palette = "Dark2") +
  theme(legend.position = "bottom") +
  xlab("D") +
  ylab("Reject")
```

```{r pcp, fig.cap = "Parallel coordinate plots of decisions made by computer vision model, conventional tests and visual tests made by human.", eval = TRUE}
library(ggpcp)
mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
  filter(!null_lineup, !attention_check) %>%
  select(type, actual_ss, model_reject_20, reject, conventional_reject) %>%
  mutate(test = rnorm(n())) %>%
  mutate(across(c(model_reject_20, conventional_reject, reject), ~ifelse(.x, "Reject", "Not reject"))) %>%
  mutate(agree = model_reject_20 == reject) %>%
  pcp_select(c(reject, model_reject_20, conventional_reject)) %>%
  group_by(type) %>%
  pcp_scale() %>%
  ungroup() %>%
  pcp_arrange(method = "from-right") %>%
  ggplot() +
  geom_pcp(aes_pcp(), linewidth = 0.1) +
  geom_pcp_labels(aes_pcp()) +
  geom_pcp_boxes(aes_pcp(), boxwidth = 0.1) +
  facet_wrap(~type, scales = "free_y") +
  theme_light() +
  scale_x_discrete(labels = c("Visual test", "Computer vision model", "Conventional test")) +
  labs(col = "D") +
  ylab("") +
  xlab("") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.title = element_blank())
```

```{r results = "hide"}
library(VennDiagram)
mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
  filter(type == "Non-linearity") %>%
  filter(!null_lineup, !attention_check) %>%
  select(unique_lineup_id, type, actual_ss, model_reject_20, reject, conventional_reject) %>%
  mutate(across(c(model_reject_20, conventional_reject, reject), ~ifelse(.x, "Reject", "Not reject"))) %>%
  pivot_longer(c(model_reject_20, reject, conventional_reject)) %>%
  group_by(name) %>%
  summarise(ids = list(unique_lineup_id[value != "Reject"])) %>%
  pull(ids) %>%
  venn.diagram(
    category.names = c("Conventional test" , "Computer vision model" , "Visual tests"),
    filename = here::here('paper/diagram/poly_venn_diagramm.png'),
    disable.logging = TRUE,
    lwd = 1,
    col = RColorBrewer::brewer.pal(3, "Dark2"),
    fill = alpha(RColorBrewer::brewer.pal(3, "Dark2"), 0.3),
    cex = 0.5,
    fontfamily = "sans",
    cat.cex = 0.5,
    cat.default.pos = "outer",
    cat.fontfamily = "sans",
    cat.pos = c(330, 30, 180),
    cat.dist = c(0, 0, 0.03),
    cat.col = RColorBrewer::brewer.pal(3, "Dark2"),
    height = 800,
    width = 800,
    resolution = 300,
    main.fontfamily = "sans",
    main.cex = 0.6,
    main = "Non-linearity",
    main.pos = c(0.25, 1.1)
  )

mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
  filter(type != "Non-linearity") %>%
  filter(!null_lineup, !attention_check) %>%
  select(unique_lineup_id, type, actual_ss, model_reject_20, reject, conventional_reject) %>%
  mutate(across(c(model_reject_20, conventional_reject, reject), ~ifelse(.x, "Reject", "Not reject"))) %>%
  pivot_longer(c(model_reject_20, reject, conventional_reject)) %>%
  group_by(name) %>%
  summarise(ids = list(unique_lineup_id[value != "Reject"])) %>%
  pull(ids) %>%
  venn.diagram(
    category.names = c("Conventional test" , "Computer vision model" , "Visual tests"),
    filename = here::here('paper/diagram/heter_venn_diagramm.png'),
    disable.logging = TRUE,
    lwd = 1,
    col = RColorBrewer::brewer.pal(3, "Dark2"),
    fill = alpha(RColorBrewer::brewer.pal(3, "Dark2"), 0.3),
    cex = 0.5,
    fontfamily = "sans",
    cat.cex = 0.5,
    cat.default.pos = "outer",
    cat.fontfamily = "sans",
    cat.pos = c(330, 30, 180),
    cat.dist = c(0, 0, 0.03),
    cat.col = RColorBrewer::brewer.pal(3, "Dark2"),
    height = 800,
    width = 800,
    resolution = 300,
    main.fontfamily = "sans",
    main.cex = 0.6,
    main = "Heteroskedasticity",
    main.pos = c(0.25, 1.1)
  )
```


```{r venn, fig.pos = "!h", fig.cap = "Venn diagram for non-rejection decisions made by conventional tests, computer vision model, and visual tests conducted by human.", eval = FALSE}
poly_venn <- magick::image_read(here::here("paper/diagram/poly_venn_diagramm.png")) %>%
  magick::image_ggplot()
heter_venn <- magick::image_read(here::here("paper/diagram/heter_venn_diagramm.png")) %>%
  magick::image_ggplot()
patchwork::wrap_plots(poly_venn, heter_venn, ncol = 2)
```


```{r eval = FALSE}
mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
  filter(!null_lineup, !attention_check) %>%
  pivot_longer(c(model_reject_20, reject, conventional_reject)) %>%
  mutate(name = ifelse(name == "model_reject_20", "Computer vision model", name)) %>%
  mutate(name = ifelse(name == "reject", "Visual test", name)) %>%
  mutate(name = ifelse(name == "conventional_reject", "Conventional test", name)) %>%
  ggplot() +
  ggbeeswarm::geom_quasirandom(aes(actual_ss, value, col = name), size = 0.5, alpha = 0.4) +
  geom_line(data = model_glm_pred_poly, aes(d, value + 1, col = "Computer vision model"), linewidth = 0.8) +
  geom_line(data = model_glm_pred_heter, aes(d, value + 1, col = "Computer vision model"), linewidth = 0.8) +
  geom_line(data = visual_glm_pred_poly, aes(d, value + 1, col = "Visual test"), linewidth = 0.8) +
  geom_line(data = visual_glm_pred_heter, aes(d, value + 1, col = "Visual test"), linewidth = 0.8) +
  geom_line(data = conv_glm_pred_poly, aes(d, value + 1, col = "Conventional test"), linewidth = 0.8) +
  geom_line(data = conv_glm_pred_heter, aes(d, value + 1, col = "Conventional test"), linewidth = 0.8) +
  geom_line(data = mutate(model_glm_pred_poly, name = "Visual test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
  geom_line(data = mutate(model_glm_pred_heter, name = "Visual test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
  geom_line(data = mutate(model_glm_pred_poly, name = "Conventional test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
  geom_line(data = mutate(model_glm_pred_heter, name = "Conventional test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
  geom_line(data = mutate(visual_glm_pred_poly, name = "Computer vision model"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
  geom_line(data = mutate(visual_glm_pred_heter, name = "Computer vision model"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
  geom_line(data = mutate(visual_glm_pred_poly, name = "Conventional test"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
  geom_line(data = mutate(visual_glm_pred_heter, name = "Conventional test"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
  geom_line(data = mutate(conv_glm_pred_poly, name = "Computer vision model"), aes(d, value + 1, col = "Conventional test"), linetype = 2, linewidth = 0.8) +
  geom_line(data = mutate(conv_glm_pred_heter, name = "Computer vision model"), aes(d, value + 1, col = "Conventional test"), linetype = 2, linewidth = 0.8) +
  geom_line(data = mutate(conv_glm_pred_poly, name = "Visual test"), aes(d, value + 1, col = "Conventional test"), linetype = 2, linewidth = 0.8) +
  geom_line(data = mutate(conv_glm_pred_heter, name = "Visual test"), aes(d, value + 1, col = "Conventional test"), linetype = 2, linewidth = 0.8) +
  facet_grid(name ~ type, scales = "free_x") +
  theme_light() +
  scale_color_brewer("", palette = "Dark2") +
  theme(legend.position = "bottom") +
  xlab("D") +
  ylab("Reject")


```


```{r delta, fig.pos = "!h", fig.cap = "A weighted detection rate vs $\\delta$-differnence plot."}
experiment %>%
  left_join(lineup_vss %>%
              group_by(unique_lineup_id) %>%
              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
  ggplot() +
  geom_point(aes(delta_diff, prop_detect), alpha = 0.3) +
  geom_smooth(aes(delta_diff, prop_detect), se = FALSE) +
  ylab("Weighted detection rate") +
  xlab(expression(delta-difference)) +
  theme_light()
```


```{r fig.pos = "!h", fig.cap = "A weighted detection rate vs $\\gamma$-number of extreme nulls plot.", eval = FALSE}
experiment %>%
  left_join(lineup_vss %>%
              group_by(unique_lineup_id) %>%
              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
  ggplot() +
  geom_point(aes(gamma_diff, prop_detect)) +
  ylab("Weighted detection rate") +
  xlab(expression(gamma-number~of~extreme~nulls)) +
  theme_light()
```


# Examples

<!-- - explain the purpose of each case here as a summary -->
<!-- - left-triangle: conventional reject, other two do not reject -->
<!-- - Boston: all three reject -->
<!-- - datasaurus: conv not reject -->

<!-- - we don't have one conv reject, model reject, human not -->

<!-- conv: reject 1, not 0 -->
<!-- model: reject 1, not 0 -->
<!-- human: reject 1, not 0 -->

<!-- left:       100 -->
<!-- boston:     111 -->
<!-- datasaurus: 011 -->

<!-- example?:   001 (probably because of outliers) -->
<!-- try "baker" dataset -->


In this section, we will utilize the trained computer vision model on three example datasets. These include the dataset associated with the residual plot displaying a "left-triangle" shape, as displayed in Figure \ref{fig:false-finding}, along with the Boston housing dataset, and the "dino" datasets from the `datasaurus` R package. 

The first example illustrates a scenario where both the computer vision model and human visual inspection successfully avoid rejecting $H_0$ when $H_0$ is true, contrary to conventional tests. This underscores the necessity of visually examining the residual plot.

In the second example, we encounter a more pronounced violation of the model, resulting in rejection of $H_0$ by all three tests. This highlights the practicality of the computer vision model, particularly for less intricate tasks.

The third example presents a situation where the model deviation is non-typical. Here, the computer vision model and human visual inspection reject $H_0$, whereas conventional tests do not. This emphasizes the benefits of visual inspection and the unique advantage of the computer vision model, which, like humans, makes decisions based on visual discoveries.

Additionally, we will outline the workflow for employing this model with the `autovi` R package.

## Left-triangle

In Section \ref{introduction}, we presented an example residual plot showcased in Figure \ref{fig:false-finding}, illustrating how humans might misinterpret the "left-triangle" shape as indicative of heteroskedasticity. Additionally, the Breusch-Pagan test yielded a rejection with a $p$-value of 0.046, despite the residuals originating from a correctly specified model. Figure \ref{fig:false-lineup} offers a lineup for this fitted model, showcasing various degrees of "left-triangle" shape across all residual plots. This phenomenon is evidently caused by the skewed distribution of the fitted values. Notably, if the residual plot were evaluated through a visual test, it would not be rejected since the true residual plot positioned at 10 can not be distinguished from the others.

Moving forward, we will employ the trained computer vision model to assess the residual plot with the assistance of the `autovi` R package. This package is accessible on the GitHub repository `TengMCing/autovi`, and users can install it by executing the command `remotes::install_github("TengMCing/autovi")` in R.

The results of the `list_keras_model()` function provide a list of available trained computer vision models. To perform the residual plot assessment, users need to supply a fitted linear regression model and select a trained Keras model to initialize the checker.

```r
checker <- auto_vi(fitted_mod = mod, 
                   keras_mod = get_keras_model("vss_phn_32"))
```


When invoking the `check()` method to execute the assessment, users can specify the number of null plots and bootstrapped samples using the `null_draws` and `boot_draws` arguments, respectively. The diagnostic results can be viewed by directly printing the object. Alternatively, a summary plot of the assessment can be generated using the `summary_plot()` method.


```r
checker$check(null_draws = 200L, boot_draws = 200L)
checker$summary_plot()
```

Figure \ref{fig:false-check} presents the results of the assessment. Notably, the observed visual signal strength is considerably lower than the 95% sample quantile of the null distribution. Moreover, the bootstrapped distribution suggests that it is highly improbable for the fitted model to be misspecified as the majority of bootstrapped fitted models will not be rejected. Thus, for this particular fitted model, both the visual test and the computer vision model will not reject $H_0$. However, the Breusch-Pagan test will reject $H_0$ because it can not effectively utilize information from null plots.

The attention map at plot B of Figure \ref{fig:false-check} suggests that the estimation is highly influenced by the top-right and bottom-right part of the residual plot, as it forms two vertices of the triangular shape. A principle component analysis is also performed on the output of the global pooling layer of the computer vision model. As mentioned in @simonyan2014very, a computer vision model built upon the convolutional blocks can be viewed as a feature extractor. For the $32 \times 32$ model, there are 256 features outputted from the global pooling layer, which can be further used for different visual tasks not limited to distance prediction. To see if these features can be effectively used for distinguishing null plots and true residual plot, we linearly project them into the PC1 vs PC2 space as shown in Figure \ref{fig:false-check}. It can be observed that because the bootstrapped plots are mostly similar to the null plots, the points drawn in different colours are mixed together. The true residual plot is also covered by both the cluster of null plots and cluster of bootstrapped plots. This accurately reflects our understanding of Figure \ref{fig:false-lineup}. 

```{r}
set.seed(452)
ori_x <- rand_lognormal()
dat <- heter_model(b = 0, x = closed_form(~-ori_x))$gen(300)

mod <- lm(y ~ x, data = dat)
my_vi <- autovi::auto_vi(fitted_mod = mod)
```

```{r}
if (!file.exists(here::here("paper/data/fig1_check.rds"))) {
  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
  my_vi$check(null_draws = 200L, boot_draws = 200L, extract_feature_from_layer = "global_max_pooling2d")
  saveRDS(my_vi$check_result, here::here("paper/data/fig1_check.rds"))
}

```

```{r}
my_vi$check_result <- readRDS(here::here("paper/data/fig1_check.rds"))
```

```{r}
define_gradient <- function() {
  reticulate::py_run_string(glue::glue(
"
import tensorflow as tf
import PIL
import numpy as np

def get_smooth_gradient(keras_mod, plot_path, target_size, input_auxiliary = None, noise_level = 0.1, n = 50):
    im = PIL.Image.open(plot_path)
    im = im.resize(target_size)
    input_im = tf.keras.utils.img_to_array(im)
    input_im = np.reshape(input_im, tuple([1]) + tuple(target_size) + tuple([3]))
    if input_auxiliary is not None:
        input_auxiliary = np.reshape(np.array(input_auxiliary), (1, 5))
    
    if input_auxiliary is not None:
        input_auxiliary = tf.Variable(input_auxiliary)
    
    im_grad = None
    sigma = noise_level * (np.max(input_im) - np.min(input_im))
    
    for i in range(n):    
        new_input_im = input_im + np.random.normal(0, sigma, input_im.shape)
        new_input_im = tf.Variable(new_input_im)
        
        with tf.GradientTape() as tape:
            if input_auxiliary is not None:
                pred = keras_mod([new_input_im, input_auxiliary])
            else:
                pred = keras_mod(new_input_im)
        new_im_grad = tape.gradient(pred, new_input_im)
        new_im_grad = tf.image.rgb_to_grayscale(new_im_grad)[0].numpy()
        if im_grad is None:
            im_grad = new_im_grad
        else:
            im_grad = im_grad + new_im_grad
    
    im_grad = im_grad / n
    
    input_im = tf.Variable(input_im)
    if input_auxiliary is not None:
        with tf.GradientTape() as tape:
            pred = keras_mod([input_im, input_auxiliary])
        auxiliary_grad = tape.gradient(pred, input_auxiliary)
    if input_auxiliary is not None:
        return (im_grad, auxiliary_grad[0].numpy())
    else:
        return im_grad
        

def get_raw_gradient(keras_mod, plot_path, target_size, input_auxiliary = None):
    im = PIL.Image.open(plot_path)
    im = im.resize(target_size)
    input_im = tf.keras.utils.img_to_array(im)
    input_im = np.reshape(input_im, tuple([1]) + tuple(target_size) + tuple([3]))
    if input_auxiliary is not None:
        input_auxiliary = np.reshape(np.array(input_auxiliary), (1, 5))
    
    input_im = tf.Variable(input_im)
    if input_auxiliary is not None:
        input_auxiliary = tf.Variable(input_auxiliary)
    
    with tf.GradientTape() as tape:
        if input_auxiliary is not None:
            pred = keras_mod([input_im, input_auxiliary])
        else:
            pred = keras_mod(input_im)
    im_grad = tape.gradient(pred, input_im)
    
    if input_auxiliary is not None:
        with tf.GradientTape() as tape:
            pred = keras_mod([input_im, input_auxiliary])
        auxiliary_grad = tape.gradient(pred, input_auxiliary)
    
    im_grad = tf.image.rgb_to_grayscale(im_grad)
    if input_auxiliary is not None:
        return (im_grad[0].numpy(), auxiliary_grad[0].numpy())
    else:
        return im_grad[0].numpy()
"
))
}

```



```{r}
if (!file.exists(here::here("paper/data/fig1_attention_map.rds"))) {
  define_gradient()
  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
  plot_path <- my_vi$plot_resid() %>%
  my_vi$save_plot()
  input_auxiliary <- my_vi$auxiliary()

  reticulate::py$get_smooth_gradient(keras_mod = keras_mod, 
                                     plot_path = plot_path, 
                                     target_size = c(32L, 32L), 
                                     input_auxiliary = input_auxiliary,
                                     noise_level = 0.5,
                                     n = 1000L) -> smooth_g
  
  x_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$x.range
  y_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$y.range
  
  smooth_att_map <- smooth_g[[1]] |> 
    as.data.frame() |>
    dplyr::mutate(row = rev(1:32)) |>
    tidyr::pivot_longer(V1:V32, names_to = "column", values_to = "gradient") |>
    dplyr::mutate(column = as.integer(gsub("V", "", column))) |>
    dplyr::mutate(column = x_range[1] + column * (x_range[2] - x_range[1])/(32 - 1)) |>
    dplyr::mutate(row = y_range[1] + row * (y_range[2] - y_range[1])/(32 - 1))
  
  scale_zero_one <- function(x) (x - min(x))/(max(x) - min(x))
  
  p <- ggplot() +
    ggplot2::geom_raster(data = mutate(smooth_att_map, gradient = scale_zero_one(gradient)), ggplot2::aes(column, row, fill = gradient), alpha = 0.7) +
    scale_fill_gradient(low = "black", high = "white") +
    ggplot2::theme_void() +
    theme(legend.position = "none")
  
  saveRDS(p, here::here("paper/data/fig1_attention_map.rds"))
  
}
```


```{r false-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots. (A) The true residual plot exhibiting "left-triangle" shape. (B) The attention map produced by computing the gradient of the output with respect to the greyscale input.  (C) The density plot of estimated distnace for null plots and bootstrapped plots. The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots. The fitted model will not be rejected since $\\hat{D} < Q_{null}(0.95)$. (D) PC2 vs PC1 plot for features extracted from the global pooling layer of the computer vision model.  '}
p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot") 
p2 <- readRDS(here::here("paper/data/fig1_attention_map.rds")) + ggtitle("(B) Attention map")
p3 <- my_vi$summary_plot() +
  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("Conventional p-value = {format(HETER_MODEL$test(dat)$p_value, digits = 3)}")) +
  theme(legend.position = "bottom", legend.box = "vertical") +
  labs(linetype = "") +
  xlab(expression(hat(D))) +
  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D))))
p4 <- my_vi$feature_pc() %>%
  mutate(type = ifelse(type == "boot", "Boot", type)) %>%
  mutate(type = ifelse(type == "null", "Null", type)) %>%
  ggplot() +
  geom_point(data = ~filter(.x, type != "observed"), aes(PC1, PC2, col = type, shape = "Simulated"), size = 2, alpha = 0.3) +
  geom_point(data = ~filter(.x, type == "observed"), aes(PC1, PC2, shape = "Observed"), size = 4, col = "black", fill = "black") +
  theme_light() +
  scale_color_brewer(palette = "Dark2") +
  scale_shape_manual(values = c("Simulated" = 19, "Observed" = 17)) +
  xlab("PC1") +
  ylab("PC2") +
  labs(col = "", shape = "") +
  theme(legend.position = "bottom", legend.box="vertical") +
  ggtitle("(D) PCA for extracted features")

patchwork::wrap_plots(p1, p2, p3, p4, nrow = 2, widths = c(1, 1), heights = c(1, 1))
```

```{r false-lineup, fig.pos = "!h", fig.cap = 'A lineup of residual plots displaying "left-triangle" visual patterns. The true residual plot occupies position 10, yet there are no discernible visual patterns that distinguish it from the other plots.'}
pos <- 10
result <- tibble()
for (i in c(1:20)[-pos]) {
  result <- bind_rows(result,
                      my_vi$rotate_resid() %>%
                        mutate(k = i))
}
result <- bind_rows(my_vi$get_fitted_and_resid() %>% 
                      mutate(k = pos),
                    result)
result %>%
  VI_MODEL$plot_lineup(remove_grid_line = TRUE,
                       remove_axis = TRUE,
                       theme = theme_light())
```

## Boston housing

The Boston housing dataset, originally published by @harrison1978hedonic, offers insights into housing in the Boston, Massachusetts area. For illustration purposes, we will utilize a reduced version from Kaggle, comprising 489 rows and 4 columns: average number of rooms per dwelling (RM), percentage of lower status of the population (LSTAT), pupil-teacher ratio by town (PTRATIO), and Median value of owner-occupied homes in $1000's (MEDV). In our analysis, MEDV will serve as the response variable, while the other columns will function as predictors in a linear regression model. Our primary focus is to detect non-linearity, because the relationships between RM and MEDV or LSTAT and MEDV are non-linear.

Figure \ref{fig:boston-check} displays the residual plot and the assessment conducted by the computer vision model. A clear non-linearity pattern resembling a "U" shape is shown in the residual plot. Furthermore, the RESET test yields a very small $p$-value. The approximated distance $\hat{D}$ significantly exceeds $Q_{null}(0.95)$, leading to rejection of $H_0$. The bootstrapped distribution also suggests that almost all the bootstrapped fitted models will be rejected, indicating that the fitted model is unlikely to be correctly specified.

Additionally, the lineup of this fitted model presented in Figure \ref{fig:boston-lineup} underscores that the "U" shape is a distinctive characteristic of the true residual plot, absent in any of the null plots. Consequently, if a visual test is conducted, $H_0$ will also be rejected.


```{r}
housing <- read_csv(here::here("paper/data/housing.csv"))
mod <- lm(MEDV ~ ., data = housing)

my_vi <- autovi::auto_vi(fitted_mod = mod)
```

```{r}
if (!file.exists(here::here("paper/data/boston_check.rds"))) {
  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
  my_vi$check(null_draws = 200L, boot_draws = 200L, extract_feature_from_layer = "global_max_pooling2d")
  saveRDS(my_vi$check_result, here::here("paper/data/boston_check.rds"))
}

```

```{r}
my_vi$check_result <- readRDS(here::here("paper/data/boston_check.rds"))
```


```{r}
if (!file.exists(here::here("paper/data/boston_attention_map.rds"))) {
  define_gradient()
  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
  plot_path <- my_vi$plot_resid() %>%
  my_vi$save_plot()
  input_auxiliary <- my_vi$auxiliary()

  reticulate::py$get_smooth_gradient(keras_mod = keras_mod, 
                                     plot_path = plot_path, 
                                     target_size = c(32L, 32L), 
                                     input_auxiliary = input_auxiliary,
                                     noise_level = 0.5,
                                     n = 1000L) -> smooth_g
  
  x_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$x.range
  y_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$y.range
  
  smooth_att_map <- smooth_g[[1]] |> 
    as.data.frame() |>
    dplyr::mutate(row = rev(1:32)) |>
    tidyr::pivot_longer(V1:V32, names_to = "column", values_to = "gradient") |>
    dplyr::mutate(column = as.integer(gsub("V", "", column))) |>
    dplyr::mutate(column = x_range[1] + column * (x_range[2] - x_range[1])/(32 - 1)) |>
    dplyr::mutate(row = y_range[1] + row * (y_range[2] - y_range[1])/(32 - 1))
  
  scale_zero_one <- function(x) (x - min(x))/(max(x) - min(x))
  
  p <- ggplot() +
    ggplot2::geom_raster(data = mutate(smooth_att_map, gradient = scale_zero_one(gradient)), ggplot2::aes(column, row, fill = gradient), alpha = 0.7) +
    scale_fill_gradient(low = "black", high = "white") +
    ggplot2::theme_void() +
    theme(legend.position = "none")
  
  saveRDS(p, here::here("paper/data/boston_attention_map.rds"))
  
}
```


```{r boston-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots. On the left, the residual plot of the Boston housing fitted model is displayed. The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots. The fitted model will be rejected since $\\hat{D} >= Q_{null}(0.95)$.'}
p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot") 
p2 <- readRDS(here::here("paper/data/boston_attention_map.rds")) + ggtitle("(B) Attention map")
p3 <- my_vi$summary_plot() +
  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("Conventional p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}")) +
  theme(legend.position = "bottom", legend.box = "vertical") +
  labs(linetype = "") +
  xlab(expression(hat(D))) +
  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D))))
p4 <- my_vi$feature_pc() %>%
  mutate(type = ifelse(type == "boot", "Boot", type)) %>%
  mutate(type = ifelse(type == "null", "Null", type)) %>%
  ggplot() +
  geom_point(data = ~filter(.x, type != "observed"), aes(PC1, PC2, col = type, shape = "Simulated"), size = 2, alpha = 0.3) +
  geom_point(data = ~filter(.x, type == "observed"), aes(PC1, PC2, shape = "Observed"), size = 4, col = "black", fill = "black") +
  theme_light() +
  scale_color_brewer(palette = "Dark2") +
  scale_shape_manual(values = c("Simulated" = 19, "Observed" = 17)) +
  xlab("PC1") +
  ylab("PC2") +
  labs(col = "", shape = "") +
  theme(legend.position = "bottom", legend.box="vertical") +
  ggtitle("(D) PCA for extracted features")

patchwork::wrap_plots(p1, p2, p3, p4, nrow = 2, widths = c(1, 1), heights = c(1, 1))
```

```{r boston-lineup, fig.pos = "!h", fig.cap = 'A lineup of residual plots for the Boston housing fitted model. The true residual plot is at position 7. It can be easily identified as the most different plot.'}
pos <- 7
result <- tibble()
for (i in c(1:20)[-pos]) {
  result <- bind_rows(result,
                      my_vi$rotate_resid() %>%
                        mutate(k = i))
}
result <- bind_rows(my_vi$get_fitted_and_resid() %>% 
                      mutate(k = pos),
                    result)
result %>%
  VI_MODEL$plot_lineup(remove_grid_line = TRUE,
                       remove_axis = TRUE,
                       theme = theme_light())
```




## Datasaurus

The computer vision model possesses the capability to detect not only typical issues like non-linearity, heteroskedasticity, and non-normality but also artifact visual patterns resembling real-world objects, as long as they do not appear in null plots. These visual patterns can be challenging to categorize in terms of model violations. Therefore, we will employ the RESET test, the Breusch-Pagan test, and the Shapiro-Wilk test [@shapiro1965analysis] for comparison.

The "dino" dataset within the `datasaurus` R package exemplifies this scenario. With only two columns, x and y, fitting a regression model to this data yields a residual plot resembling a "dinosaur," as displayed in Figure \ref{fig:dino-check}. Unsurprisingly, this distinct residual plot stands out in a lineup, as shown in Figure \ref{fig:dino-lineup}. A visual test conducted by humans would undoubtedly reject $H_0$.

According to the residual plot assessment by the computer vision model, $\hat{D}$ exceeds $Q_{null}(0.95)$, warranting a rejection of $H_0$. Additionally, most of the bootstrapped fitted models will be rejected, indicating an misspecified model. However, both the RESET test and the Breusch-Pagan test yield $p$-values greater than 0.3, leading to a non-rejection of $H_0$. Only the Shapiro-Wilk test rejects the normality assumption with a small $p$-value.

In practice, without accessing the residual plot, it would be challenging to identify the artificial pattern of the residuals. Moreover, conducting a normality test for a fitted regression model is not always standard practice among analysts. Even when performed, violating the normality assumption is sometimes deemed acceptable, especially considering the application of quasi-maximum likelihood estimation in linear regression. This example underscores the importance of evaluating residual plots and highlights how the proposed computer vision model can facilitate this process.


```{r}
dino <- datasauRus::datasaurus_dozen %>% filter(dataset == "dino")
mod <- lm(y ~ ., data = select(dino, -dataset))

my_vi <- autovi::auto_vi(fitted_mod = mod)
```

```{r}
if (!file.exists(here::here("paper/data/dino_check.rds"))) {
  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
  my_vi$check(null_draws = 200L, boot_draws = 200L, extract_feature_from_layer = "global_max_pooling2d")
  saveRDS(my_vi$check_result, here::here("paper/data/dino_check.rds"))
}
```

```{r}
my_vi$check_result <- readRDS(here::here("paper/data/dino_check.rds"))
```

```{r}
if (!file.exists(here::here("paper/data/dino_attention_map.rds"))) {
  define_gradient()
  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
  plot_path <- my_vi$plot_resid() %>%
  my_vi$save_plot()
  input_auxiliary <- my_vi$auxiliary()

  reticulate::py$get_smooth_gradient(keras_mod = keras_mod, 
                                     plot_path = plot_path, 
                                     target_size = c(32L, 32L), 
                                     input_auxiliary = input_auxiliary,
                                     noise_level = 0.5,
                                     n = 1000L) -> smooth_g
  
  x_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$x.range
  y_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$y.range
  
  smooth_att_map <- smooth_g[[1]] |> 
    as.data.frame() |>
    dplyr::mutate(row = rev(1:32)) |>
    tidyr::pivot_longer(V1:V32, names_to = "column", values_to = "gradient") |>
    dplyr::mutate(column = as.integer(gsub("V", "", column))) |>
    dplyr::mutate(column = x_range[1] + column * (x_range[2] - x_range[1])/(32 - 1)) |>
    dplyr::mutate(row = y_range[1] + row * (y_range[2] - y_range[1])/(32 - 1))
  
  scale_zero_one <- function(x) (x - min(x))/(max(x) - min(x))
  
  p <- ggplot() +
    ggplot2::geom_raster(data = mutate(smooth_att_map, gradient = scale_zero_one(gradient)), ggplot2::aes(column, row, fill = gradient), alpha = 0.7) +
    scale_fill_gradient(low = "black", high = "white") +
    ggplot2::theme_void() +
    theme(legend.position = "none")
  
  saveRDS(p, here::here("paper/data/dino_attention_map.rds"))
  
}
```

```{r dino-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots. On the left, the residual plot exhitbiting a "dinosaur" shape is displayed. The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots. The fitted model will be rejected since $\\hat{D} >= Q_{null}(0.95)$.'}

p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot")
p2 <- readRDS(here::here("paper/data/dino_attention_map.rds")) + ggtitle("(B) Attention map")
p3 <- my_vi$summary_plot() +
  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("RESET test p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}\n Breusch-Pagan test p-value = {format(lmtest::bptest(mod)$p.value, digits = 3)} \n Shapiro-Wilk test p-value = {format(shapiro.test(mod$residuals)$p.value, digits = 3)}")) +
  theme(legend.position = "bottom", legend.box = "vertical") +
  labs(linetype = "") +
  xlab(expression(hat(D))) +
  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D))))
p4 <- my_vi$feature_pc() %>%
  mutate(type = ifelse(type == "boot", "Boot", type)) %>%
  mutate(type = ifelse(type == "null", "Null", type)) %>%
  ggplot() +
  geom_point(data = ~filter(.x, type != "observed"), aes(PC1, PC2, col = type, shape = "Simulated"), size = 2, alpha = 0.3) +
  geom_point(data = ~filter(.x, type == "observed"), aes(PC1, PC2, shape = "Observed"), size = 4, col = "black", fill = "black") +
  theme_light() +
  scale_color_brewer(palette = "Dark2") +
  scale_shape_manual(values = c("Simulated" = 19, "Observed" = 17)) +
  xlab("PC1") +
  ylab("PC2") +
  labs(col = "", shape = "") +
  theme(legend.position = "bottom", legend.box="vertical") +
  ggtitle("(D) PCA for extracted features")

patchwork::wrap_plots(p1, p2, p3, p4, nrow = 2, widths = c(1, 1), heights = c(1, 1))
```

```{r dino-lineup, fig.pos = "!h", fig.cap = 'A lineup of residual plots for the fitted model on the "dinosaur" dataset. The true residual plot is at position 17. It can be easily identified as the most different plot as the visual pattern is extremly artificial.'}
pos <- 17
result <- tibble()
for (i in c(1:20)[-pos]) {
  result <- bind_rows(result,
                      my_vi$rotate_resid() %>%
                        mutate(k = i))
}
result <- bind_rows(my_vi$get_fitted_and_resid() %>% 
                      mutate(k = pos),
                    result)
result %>%
  VI_MODEL$plot_lineup(remove_grid_line = TRUE,
                       remove_axis = TRUE,
                       theme = theme_light())
```

<!-- ## When the model works -->

<!-- - simple examples (non-linearity, heteroskedasticity, ...) -->
<!-- - datasaurus -->

<!-- ## When the model does not work -->

<!-- - human detect but model does not -->
<!-- - cartoon residuals? -->


<!-- ## Workflow: how one use this model? (small showcase)  -->

<!-- # Discussion -->

<!-- There are other kinds of residual departures like autocorrelation that are not considered in this study. The primary goal of this study is to establish a new way of evaluating residual plot and conducting visual test with computer vision models. Building computer vision models for other model violations and other types of diagnostic plots could be future directions of this field.    -->

# Extension to other types of regression model

1. Use raw residuals.
  - D (effect size) for null plots will be greater than zero.
  - Can not assume D for actual plots are greater than D for null plots any more. -> KL null - normal, KL actual - normal
  - Need to use a two-sided test instead of a one-sided test.
  - It may fail completely due to the identification issue.
2. Use transformed residuals that roughly follow a multivariate normal distribution.
  - The one-sided test will still work. 
  - Need to ensure that the transformed actual plot will have residual distribution different from the transformed null distribution. 
3. Train a new model.
  - KL divergence may not be the appropriate measure for the difference between the residual distributions. Could consider using a new distance measure.



# Conclusion

- Brief discussion
- Summary of findings
- Contributions to the field

In this study, we have introduced a distance metric based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted linear regression model and the reference residual distribution assumed under correct model specification. This distance metric effectively captures the magnitude of model violations in misspecified models. We propose a computer vision model to approximate this distance, utilizing the residual plot of the fitted model as input. The resulting approximated distance serves as the foundation for constructing a single Model Violation Index (MVI), facilitating the quantification of various model violations.

Moreover, the approximated distance enables the development of a formal statistical testing procedure by evaluating a large number of null plots generated from the fitted model. Additionally, employing bootstrapping techniques and refitting the regression model allows us to ascertain how frequently the fitted model is considered misspecified if data were repeatedly obtained from the same data generating process.

The trained computer vision model demonstrates strong performance on both the training and test sets, although it exhibits slightly lower performance on residual plots with non-linearity visual patterns compared to other types of violations. The statistical tests relying on the approximated distance predicted by the computer vision model exhibit lower sensitivity compared to conventional tests but higher sensitivity compared to visual tests conducted by humans. While the approximated distance generally mirrors the strength of the visual signal perceived by humans, there remains scope for further improvement in its performance.

Several examples are provided to showcase the effectiveness of the proposed method across different scenarios, emphasizing the similarity between visual tests and distance-based tests. Overall, both visual tests and distance-based tests can be viewed as combined tests, aiming to assess the correctness of all model assumptions collectively. In contrast, individual residual diagnostic tests such as the RESET test and the Breusch-Pagan test only evaluate specific model assumptions.

In practice, selecting an appropriate set of statistical tests for regression diagnostics can be challenging, particularly given the necessity of adjusting the significance level for each test. 

Our method holds significant value as it helps alleviate a portion of analysts' workload associated with assessing residual plots. While we recommend analysts to continue reading residual plots whenever feasible, as they offer invaluable insights, our approach serves as a valuable tool for automating the diagnostic process or for supplementary purposes when needed.


- Future directions



# Acknowledgements {-}

- Packages used
- Website link


```{r eval = FALSE}

# This shows that even if the model can 100% accurately predict the distance for not null plots, 
# the statistical tests based on approximated distance will still reject more cases than visual tests
# So it has very little to do with the accuracy of the prediction
experiment %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(model_reject_true = actual_ss > vss_null_95) %>%
  filter(model_reject_true != reject) %>%
  ggplot() +
  ggbeeswarm::geom_quasirandom(aes(actual_ss, reject)) +
  ggtitle("Unmatched cases") +
  ylab("Visual tests reject") +
  xlab("D")
```

```{r eval = FALSE}

map_dbl(10^seq(-10, 0, 0.01), function(crt) {
  experiment %>%
    filter(!null_lineup) %>%
    filter(!attention_check) %>%
    mutate(mvi = vss - log(n) + 10) %>%
    mutate(actual_ss_cut = conventional_p_value <= crt) %>%
    mutate(agree = actual_ss_cut == reject) %>%
    pull(agree) %>%
    mean()
}) %>%
  max()

experiment %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(mvi = vss - log(n) + 10) %>%
  ggplot() +
  ggbeeswarm::geom_quasirandom(aes(vss, reject)) +
  ggtitle("Unmatched cases") +
  ylab("Visual tests reject") +
  xlab("D")
  mutate(model_reject_true = actual_ss > vss_null_95) %>%
  filter(model_reject_true != reject)
```



